---
title: "Nutrient Limitation in the MDV of Antarctica"
subtitle: "  \n Surface Growth - Counts (Point-Intercept)"
author: "Adam J. Solon"
date: "`r Sys.Date()`"
#output: html_document
output: github_document
#output: 
  #pdf_document:
    #includes:  
     # in_header: my_header.txt
    #toc: TRUE
    #fig_width: 8
    #fig_height: 6
    #fig_caption: true
 #fontsize: 12pt
#editor_options: 
#  chunk_output_type: console
---

## Script Summary  
This report quantifies surface growth of primary producers from MDV soils with the point-intercept method under a microscope with 45x magnification (i.e., following a random grid-like pattern, every time the reticle is touching a primary producer it counts a '1' this is repeated 100x for a percent cover estimate).  The experiment involved soils from Taylor Valley in the McMurdo Dry Valleys (MDV) of Antarctica and were collected from mountain slopes in one of three lake basins: Bonney(interior), Hoare(intermediate), and Fryxell(coastal).  There are five groups from each basin: Bulk (homogenized, starting soils), CTRL (controls with H2O only), N (nitrogen + H2O only), P (phosphorus + H2O only), and NP (N and P combined + H2O). 

### Steps of this pipeline:  
1.  Create and organize directories
2.  Load R packages
3.  Subset according to basin
4.  Input files
5.  Format Files
6.  Calculate percent cover
7.  Save files for further analyses

```{r, echo = FALSE, include = FALSE}
# Change identifiers to your system and file naming. 
user <- "C:/Users/adams" # computer user name
folder <- "/Desktop/Projects"
project <- "/MDV_Soils_nutrient_limitation/manuscript1"
analyses <- "/analyses"
data  <- "/point_intercept"
analysis1 <- "/count_processing"
analysis2 <- "/SC_surface_cover"
analysis3 <- "/RR_response_ratio"
analysis4 <- "/GR_growth_rate"
ecosystem <- "lab microcosms - Taylor Valley soils" # Define the environment for use in report language.
project.fp <- paste0(user, folder, project)
```

## Pipeline

#### Set paths and create directories 

```{r set paths for project, echo = FALSE, include = FALSE}
# First define the project and project folders. 

# Create project directory
###################################################
project.fp <- paste0(user, folder, project)
if (!dir.exists(project.fp)) dir.create(project.fp)

# Create sub-directory for all analyses for project 
###################################################
analyses.fp <- paste0(project.fp, analyses)
if (!dir.exists(analyses.fp)) dir.create(analyses.fp)

# Create sub-directory for data type 
###################################################
data.fp <- paste0(analyses.fp, data)
if (!dir.exists(data.fp)) dir.create(data.fp)

# Create sub-directory for specific analysis 
###################################################
analysis1.fp <- paste0(data.fp, analysis1)
if (!dir.exists(analysis1.fp)) dir.create(analysis1.fp)

# Create sub-directories for analysis 
###################################################

# Create directory for 'before' pipeline inputs
input.fp <- paste0(analysis1.fp, "/input")
if (!dir.exists(input.fp)) dir.create(input.fp)

# Create directory for 'within' pipeline R objects 
objects.fp <- paste0(analysis1.fp, "/objects")
if (!dir.exists(objects.fp)) dir.create(objects.fp)

# Create directory for 'after' pipeline outputs 
output.fp <- paste0(analysis1.fp, "/output")
if (!dir.exists(output.fp)) dir.create(output.fp)

# Check for the folders here as a sanity check. Should see "Input" and "Objects" if starting from scratch.
list.files(analysis1.fp) 

```

#### Create additional sub-directories by analysis and basin  

```{r echo = FALSE, include = FALSE}
# Create sub-directory for Surface Cover (SC) analysis
###################################################
analysis2.fp <- paste0(data.fp, analysis2)
if (!dir.exists(analysis2.fp)) dir.create(analysis2.fp)

# Create sub-directories for SC 
###################################################

# Create directory for 'before' pipeline inputs
input2.fp <- paste0(analysis2.fp, "/input")
if (!dir.exists(input2.fp)) dir.create(input2.fp)

# Create directory for 'within' pipeline R objects 
objects2.fp <- paste0(analysis2.fp, "/objects")
if (!dir.exists(objects2.fp)) dir.create(objects2.fp)

# Create directory for 'after' pipeline outputs 
output2.fp <- paste0(analysis2.fp, "/output")
if (!dir.exists(output2.fp)) dir.create(output2.fp)

# Check for the folders here as a sanity check. Should see "Input" and "Objects" if starting from scratch.
list.files(analysis2.fp) 

# Create further sub-directories by basin
###################################################

# Create directory for 'within' pipeline R objects 
objects2.b.fp <- paste0(objects2.fp, "/bonney")
if (!dir.exists(objects2.b.fp)) dir.create(objects2.b.fp)

# Create directory for 'after' pipeline outputs 
output2.b.fp <- paste0(output2.fp, "/bonney")
if (!dir.exists(output2.b.fp)) dir.create(output2.b.fp)

# Create sub-directories for Hoare basin 
###################################################

# Create directory for 'within' pipeline R objects 
objects2.h.fp <- paste0(objects2.fp, "/hoare")
if (!dir.exists(objects2.h.fp)) dir.create(objects2.h.fp)

# Create directory for 'after' pipeline outputs 
output2.h.fp <- paste0(output2.fp, "/hoare")
if (!dir.exists(output2.h.fp)) dir.create(output2.h.fp)

# Create sub-directories for Fryxell basin 
###################################################

# Create directory for 'within' pipeline R objects 
objects2.f.fp <- paste0(objects2.fp, "/fryxell")
if (!dir.exists(objects2.f.fp)) dir.create(objects2.f.fp)

# Create directory for 'after' pipeline outputs 
output2.f.fp <- paste0(output2.fp, "/fryxell")
if (!dir.exists(output2.f.fp)) dir.create(output2.f.fp)

```


```{r echo = FALSE, include = FALSE}
# Create sub-directory for Response Ratios (RR) analysis
###################################################
analysis3.fp <- paste0(data.fp, analysis3)
if (!dir.exists(analysis3.fp)) dir.create(analysis3.fp)

# Create sub-directories for RR 
###################################################

# Create directory for 'before' pipeline inputs
input3.fp <- paste0(analysis3.fp, "/input")
if (!dir.exists(input3.fp)) dir.create(input3.fp)

# Create directory for 'within' pipeline R objects 
objects3.fp <- paste0(analysis3.fp, "/objects")
if (!dir.exists(objects3.fp)) dir.create(objects3.fp)

# Create directory for 'after' pipeline outputs 
output3.fp <- paste0(analysis3.fp, "/output")
if (!dir.exists(output3.fp)) dir.create(output3.fp)

# Check for the folders here as a sanity check. Should see "Input" and "Objects" if starting from scratch.
list.files(analysis3.fp) 

# Create further sub-directories by basin
###################################################
# Bonney
# Create directory for 'within' pipeline R objects 
objects3.b.fp <- paste0(objects3.fp, "/bonney")
if (!dir.exists(objects3.b.fp)) dir.create(objects3.b.fp)

# Create directory for 'after' pipeline outputs 
output3.b.fp <- paste0(output3.fp, "/bonney")
if (!dir.exists(output3.b.fp)) dir.create(output3.b.fp)

# Create sub-directories for Hoare basin 
###################################################

# Create directory for 'within' pipeline R objects 
objects3.h.fp <- paste0(objects3.fp, "/hoare")
if (!dir.exists(objects3.h.fp)) dir.create(objects3.h.fp)

# Create directory for 'after' pipeline outputs 
output3.h.fp <- paste0(output3.fp, "/hoare")
if (!dir.exists(output3.h.fp)) dir.create(output3.h.fp)

# Create sub-directories for Fryxell basin 
###################################################

# Create directory for 'within' pipeline R objects 
objects3.f.fp <- paste0(objects3.fp, "/fryxell")
if (!dir.exists(objects3.f.fp)) dir.create(objects3.f.fp)

# Create directory for 'after' pipeline outputs 
output3.f.fp <- paste0(output3.fp, "/fryxell")
if (!dir.exists(output3.f.fp)) dir.create(output3.f.fp)

```

```{r echo = FALSE, include = FALSE}
# Create sub-directory for Growth Rates (GR) analysis
###################################################
analysis4.fp <- paste0(data.fp, analysis4)
if (!dir.exists(analysis4.fp)) dir.create(analysis4.fp)

# Create sub-directories for GR 
###################################################

# Create directory for 'before' pipeline inputs
input4.fp <- paste0(analysis4.fp, "/input")
if (!dir.exists(input4.fp)) dir.create(input4.fp)

# Create directory for 'within' pipeline R objects 
objects4.fp <- paste0(analysis4.fp, "/objects")
if (!dir.exists(objects4.fp)) dir.create(objects4.fp)

# Create directory for 'after' pipeline outputs 
output4.fp <- paste0(analysis4.fp, "/output")
if (!dir.exists(output4.fp)) dir.create(output4.fp)

# Check for the folders here as a sanity check. Should see "Input" and "Objects" if starting from scratch.
list.files(analysis4.fp) 

# Create further sub-directories by basin
###################################################

# Create directory for 'within' pipeline R objects 
objects4.b.fp <- paste0(objects4.fp, "/bonney")
if (!dir.exists(objects4.b.fp)) dir.create(objects4.b.fp)

# Create directory for 'after' pipeline outputs 
output4.b.fp <- paste0(output4.fp, "/bonney")
if (!dir.exists(output4.b.fp)) dir.create(output4.b.fp)

# Create sub-directories for Hoare basin 
###################################################

# Create directory for 'within' pipeline R objects 
objects4.h.fp <- paste0(objects4.fp, "/hoare")
if (!dir.exists(objects4.h.fp)) dir.create(objects4.h.fp)

# Create directory for 'after' pipeline outputs 
output4.h.fp <- paste0(output4.fp, "/hoare")
if (!dir.exists(output4.h.fp)) dir.create(output4.h.fp)

# Create sub-directories for Fryxell basin 
###################################################

# Create directory for 'within' pipeline R objects 
objects4.f.fp <- paste0(objects4.fp, "/fryxell")
if (!dir.exists(objects4.f.fp)) dir.create(objects4.f.fp)

# Create directory for 'after' pipeline outputs 
output4.f.fp <- paste0(output4.fp, "/fryxell")
if (!dir.exists(output4.f.fp)) dir.create(output4.f.fp)

```

#### Session Information  

```{r Install and load packages, echo = FALSE, include = FALSE}

# install.packages("tidyverse")
# install.packages("knitr")
# install.packages("lattice")

library(tidyverse); packageVersion("tidyverse")
library(knitr); packageVersion("knitr")
library(lattice); packageVersion("lattice")

knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(include = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

```  
  
* r version: `r getRversion()`
* RStudio version: `r rstudioapi::versionInfo()$version`
* r packages:  
  tidyverse, `r packageVersion("tidyverse")` 
  knitr, `r packageVersion("knitr")`  
  lattice, `r packageVersion("lattice")`  
  
#### Input files 
1. Count data from microcosms, filename 'MCosms_II_metadata.csv'

```{r input data files, echo = FALSE, include = FALSE}
# input data files
# Count data from point-intercept (microscopy)
counts.fp <- paste0(input.fp, "/MCosms_II_microscope_counts_all.csv") 

#input file
df <- read.csv(counts.fp)

```

#### Format files  
1. Remove NAs (i.e., the spot checks with field-of-view before growth started and point-intercept was used, or Bonney plate 3 which was lost during experiment)
2. Re-label 'N+P' as 'NP'
3. Create a factor1 (N) and factor2 (P) as boolean variables
4. Reclassify explanatory variables (factor1, factor2, and treatment) as factors with two, two, or four levels respectively
3. Subset by basin

```{r echo = FALSE, include = FALSE}
#inspect data structures of input files
#metadata file
str(df)

#remove NAs
df.1 <- df %>% drop_na()

#recode N+P treatment as NP
df.1 <- df.1 %>%
  mutate(treatment = replace(treatment, treatment == "N+P", "NP"))

#create boolean variables for each treatment in new columns
df.1$N <- ifelse(df.1$treatment %in% c("N", "NP"), TRUE, FALSE)
#
#create boolean variables for each treatment in new columns
df.1$P <- ifelse(df.1$treatment %in% c("P", "NP"), TRUE, FALSE)
#
#reclassify plate_number, n, and p, as factors
df.1 <- df.1 %>% 
  mutate_at(vars(plate_number, treatment, N, P,), factor)

#subset by basin
#bonney
df.b <- df.1 %>% filter(basin == "Bonney") 

#hoare
df.h <- df.1 %>% filter(basin == "Hoare") 

#fryxell
df.f <- df.1 %>% filter(basin == "Fryxell") 
```

#### Compute percent cover for each count column - Bonney

```{r echo = FALSE, include = FALSE}
#add up all counts across columns for 'total' surface count
df.b <- df.b %>% mutate(total = green_sphere + 
                          green_dot +
                          green_branching +
                          brown_dot + 
                          green_rope + 
                          green_patch + 
                          yellow_patch +
                          green_filaments + 
                          fungi +  
                          moss + 
                          orange_sphere,
                        .after = orange_sphere)

#add up prijmary producer counts across columns for 'primary' surface count
df.b <- df.b %>% mutate(primary = green_sphere + 
                          green_dot +
                          green_branching +
                          brown_dot + 
                          green_rope + 
                          green_patch + 
                          yellow_patch +
                          green_filaments + 
                          moss + 
                          orange_sphere,
                        .after = total)

#compute percent total cover
df.b <- df.b %>% mutate(total_cover = (total / num_counts) * 100,
                        .after = primary)

#compute percent primary producer cover
df.b <- df.b %>% mutate(primary_cover = (primary / num_counts) * 100,
                        .after = total_cover)

#compute percent primary producer cover between 0 and 1
df.b <- df.b %>% mutate(primary_cover_2 = (primary / num_counts),
                        .after = primary_cover)

#compute percent green sphere cover
df.b <- df.b %>% mutate(gs_cover = (green_sphere / num_counts) * 100,
                        .after = primary_cover_2)

#compute percent green dot cover
df.b <- df.b %>% mutate(gd_cover = (green_dot / num_counts) * 100,
                        .after = gs_cover)

#compute percent green branching cover
df.b <- df.b %>% mutate(gb_cover = (green_branching / num_counts) * 100,
                        .after = gd_cover)

#compute percent brown dot cover
df.b <- df.b %>% mutate(bd_cover = (brown_dot / num_counts) * 100,
                        .after = gb_cover)

#compute percent green rope cover
df.b <- df.b %>% mutate(gr_cover = (green_rope / num_counts) * 100,
                        .after = bd_cover)

#compute percent green patch cover
df.b <- df.b %>% mutate(gp_cover = (green_patch / num_counts) * 100,
                        .after = gr_cover)

#compute percent yellow patch cover
df.b <- df.b %>% mutate(yp_cover = (yellow_patch / num_counts) * 100,
                        .after = gp_cover)

#compute percent green filaments cover
df.b <- df.b %>% mutate(gf_cover = (green_filaments / num_counts) * 100,
                        .after = yp_cover)

#compute percent fungi cover
df.b <- df.b %>% mutate(fg_cover = (fungi / num_counts) * 100,
                        .after = gf_cover)

#compute percent moss cover
df.b <- df.b %>% mutate(ms_cover = (moss / num_counts) * 100,
                        .after = fg_cover)

#compute percent orange sphere cover
df.b <- df.b %>% mutate(os_cover = (orange_sphere / num_counts) * 100,
                        .after = ms_cover)


#create vector with days for each plate by date column
days.b <- rep(c("35", "42", "49", "56", "63", "70"), each = 19)

#join vector to data frame
df.b.1 <- cbind(df.b, days.b)

#recode days from character to integer
df.b.1$days.b <- as.integer(df.b.1$days.b)

##Save as r object and .txt file for further use
#save as an R file
saveRDS(df.b.1, paste0(objects.fp, "/df.b.1.rds"))

#save as .txt file 
write.table(df.b.1, file = paste0(output.fp, "/Bonney_counts_percent.txt"), 
            sep = "\t", row.names = FALSE, col.names = TRUE)

#save as an R file
saveRDS(df.b.1, paste0(objects2.b.fp, "/df.b.1.rds"))

#save as .txt file 
write.table(df.b.1, file = paste0(input2.fp, "/Bonney_counts_percent.txt"), 
            sep = "\t", row.names = FALSE, col.names = TRUE)

#save as an R file
saveRDS(df.b.1, paste0(objects4.b.fp, "/df.b.1.rds"))

#save as .txt file 
write.table(df.b.1, file = paste0(input4.fp, "/Bonney_counts_percent.txt"), 
            sep = "\t", row.names = FALSE, col.names = TRUE)
```

#### Compute percent cover for each count column - Hoare

```{r echo = FALSE, include = FALSE}
#add up all counts across columns for 'total' surface count
df.h <- df.h %>% mutate(total = green_sphere + 
                          green_dot +
                          green_branching +
                          brown_dot + 
                          green_rope + 
                          green_patch + 
                          yellow_patch +
                          green_filaments + 
                          fungi +  
                          moss + 
                          orange_sphere,
                        .after = orange_sphere)

#add up prijmary producer counts across columns for 'primary' surface count
df.h <- df.h %>% mutate(primary = green_sphere + 
                          green_dot +
                          green_branching +
                          brown_dot + 
                          green_rope + 
                          green_patch + 
                          yellow_patch +
                          green_filaments + 
                          moss + 
                          orange_sphere,
                        .after = total)

#compute percent total cover
df.h <- df.h %>% mutate(total_cover = (total / num_counts) * 100,
                        .after = primary)

#compute percent primary producer cover
df.h <- df.h %>% mutate(primary_cover = (primary / num_counts) * 100,
                        .after = total_cover)

#compute percent primary producer cover between 0 and 1
df.h <- df.h %>% mutate(primary_cover_2 = (primary / num_counts),
                        .after = primary_cover)

#compute percent green sphere cover
df.h <- df.h %>% mutate(gs_cover = (green_sphere / num_counts) * 100,
                        .after = primary_cover_2)

#compute percent green dot cover
df.h <- df.h %>% mutate(gd_cover = (green_dot / num_counts) * 100,
                        .after = gs_cover)

#compute percent green branching cover
df.h <- df.h %>% mutate(gb_cover = (green_branching / num_counts) * 100,
                        .after = gd_cover)

#compute percent brown dot cover
df.h <- df.h %>% mutate(bd_cover = (brown_dot / num_counts) * 100,
                        .after = gb_cover)

#compute percent green rope cover
df.h <- df.h %>% mutate(gr_cover = (green_rope / num_counts) * 100,
                        .after = bd_cover)

#compute percent green patch cover
df.h <- df.h %>% mutate(gp_cover = (green_patch / num_counts) * 100,
                        .after = gr_cover)

#compute percent yellow patch cover
df.h <- df.h %>% mutate(yp_cover = (yellow_patch / num_counts) * 100,
                        .after = gp_cover)

#compute percent green filaments cover
df.h <- df.h %>% mutate(gf_cover = (green_filaments / num_counts) * 100,
                        .after = yp_cover)

#compute percent fungi cover
df.h <- df.h %>% mutate(fg_cover = (fungi / num_counts) * 100,
                        .after = gf_cover)

#compute percent moss cover
df.h <- df.h %>% mutate(ms_cover = (moss / num_counts) * 100,
                        .after = fg_cover)

#compute percent orange sphere cover
df.h <- df.h %>% mutate(os_cover = (orange_sphere / num_counts) * 100,
                        .after = ms_cover)


#create vector with days for each plate by date column
days.h <- rep(c("28", "35", "42", "49", "56", "63"), each = 20)

#join vector to data frame
df.h.1 <- cbind(df.h, days.h)

#recode days from character to integer
df.h.1$days.h <- as.integer(df.h.1$days.h)

##Save as r object and .txt file for further use
#save as an R file
saveRDS(df.h.1, paste0(objects.fp, "/df.h.1.rds"))

#save as .txt file 
write.table(df.h.1, file = paste0(output.fp, "/Hoare_counts_percent.txt"), 
            sep = "\t", row.names = FALSE, col.names = TRUE)

#save as an R file
saveRDS(df.h.1, paste0(objects2.h.fp, "/df.h.1.rds"))

#save as .txt file 
write.table(df.h.1, file = paste0(input2.fp, "/Hoare_counts_percent.txt"), 
            sep = "\t", row.names = FALSE, col.names = TRUE)

#save as an R file
saveRDS(df.h.1, paste0(objects4.h.fp, "/df.h.1.rds"))

#save as .txt file 
write.table(df.h.1, file = paste0(input4.fp, "/Hoare_counts_percent.txt"), 
            sep = "\t", row.names = FALSE, col.names = TRUE)
```

#### Compute percent cover for each count column - Fryxell

```{r echo = FALSE, include = FALSE}
#add up all counts across columns for 'total' surface count
df.f <- df.f %>% mutate(total = green_sphere + 
                          green_dot +
                          green_branching +
                          brown_dot + 
                          green_rope + 
                          green_patch + 
                          yellow_patch +
                          green_filaments + 
                          fungi +  
                          moss + 
                          orange_sphere,
                        .after = orange_sphere)

#add up prijmary producer counts across columns for 'primary' surface count
df.f <- df.f %>% mutate(primary = green_sphere + 
                          green_dot +
                          green_branching +
                          brown_dot + 
                          green_rope + 
                          green_patch + 
                          yellow_patch +
                          green_filaments + 
                          moss + 
                          orange_sphere,
                        .after = total)

#compute percent total cover
df.f <- df.f %>% mutate(total_cover = (total / num_counts) * 100,
                        .after = primary)

#compute percent primary producer cover
df.f <- df.f %>% mutate(primary_cover = (primary / num_counts) * 100,
                        .after = total_cover)

#compute percent primary producer cover between 0 and 1
df.f <- df.f %>% mutate(primary_cover_2 = (primary / num_counts),
                        .after = primary_cover)

#compute percent green sphere cover
df.f <- df.f %>% mutate(gs_cover = (green_sphere / num_counts) * 100,
                        .after = primary_cover_2)

#compute percent green dot cover
df.f <- df.f %>% mutate(gd_cover = (green_dot / num_counts) * 100,
                        .after = gs_cover)

#compute percent green branching cover
df.f <- df.f %>% mutate(gb_cover = (green_branching / num_counts) * 100,
                        .after = gd_cover)

#compute percent brown dot cover
df.f <- df.f %>% mutate(bd_cover = (brown_dot / num_counts) * 100,
                        .after = gb_cover)

#compute percent green rope cover
df.f <- df.f %>% mutate(gr_cover = (green_rope / num_counts) * 100,
                        .after = bd_cover)

#compute percent green patch cover
df.f <- df.f %>% mutate(gp_cover = (green_patch / num_counts) * 100,
                        .after = gr_cover)

#compute percent yellow patch cover
df.f <- df.f %>% mutate(yp_cover = (yellow_patch / num_counts) * 100,
                        .after = gp_cover)

#compute percent green filaments cover
df.f <- df.f %>% mutate(gf_cover = (green_filaments / num_counts) * 100,
                        .after = yp_cover)

#compute percent fungi cover
df.f <- df.f %>% mutate(fg_cover = (fungi / num_counts) * 100,
                        .after = gf_cover)

#compute percent moss cover
df.f <- df.f %>% mutate(ms_cover = (moss / num_counts) * 100,
                        .after = fg_cover)

#compute percent orange sphere cover
df.f <- df.f %>% mutate(os_cover = (orange_sphere / num_counts) * 100,
                        .after = ms_cover)


#create vector with days for each plate by date column
days.f <- rep(c("28", "35", "42", "49", "56", "63"), each = 20)

#join vector to data frame
df.f.1 <- cbind(df.f, days.f)

#recode days from character to integer
df.f.1$days.f <- as.integer(df.f.1$days.f)

##Save as r object and .txt file for further use
#save as an R file
saveRDS(df.f.1, paste0(objects.fp, "/df.f.1.rds"))

#save as .txt file 
write.table(df.f.1, file = paste0(output.fp, "/Fryxell_counts_percent.txt"), 
            sep = "\t", row.names = FALSE, col.names = TRUE)

#save as an R file
saveRDS(df.f.1, paste0(objects2.f.fp, "/df.f.1.rds"))

#save as .txt file 
write.table(df.f.1, file = paste0(input2.fp, "/Fryxell_counts_percent.txt"), 
            sep = "\t", row.names = FALSE, col.names = TRUE)

#save as an R file
saveRDS(df.f.1, paste0(objects4.f.fp, "/df.f.1.rds"))

#save as .txt file 
write.table(df.f.1, file = paste0(input4.fp, "/Fryxell_counts_percent.txt"), 
            sep = "\t", row.names = FALSE, col.names = TRUE)
```

## Plots  

#### Plot growth by plate - Bonney  

```{r echo = FALSE, include = FALSE}
#Bonney
#xyplot for each plate
xyplot.b <- xyplot(primary_cover ~ days.b|plate_number, 
       data = df.b.1,
       groups = treatment, 
       xlab="Days since start of experiment",
       ylab="Surface Cover (%)",
       #main="Surface Cover by Plate- Bonney Basin Soils",
       type=c('g','p','l'),
       pch = 16, cex = 0.5,
       index.cond = function(x,y)max(y))
```
  
```{r echo = FALSE, fig.pos = 'h', fig.cap = "Plot of each microcosms surface cover over time for Bonney basin soils.  Counts begin on day 35 and end on day 70"}
plot(xyplot.b)
```

#### Plot growth by plate - Hoare  

```{r echo = FALSE, include = FALSE}
#Hoare
#xyplot for each plate
xyplot.h <- xyplot(primary_cover ~ days.h|plate_number, 
       data = df.h.1,
       groups = treatment, 
       xlab="Days since start of experiment",
       ylab="Surface Cover (%)",
       #main="Surface Cover by Plate- Bonney Basin Soils",
       type=c('g','p','l'),
       pch = 16, cex = 0.5,
       index.cond = function(x,y)max(y))

```
  
```{r echo = FALSE, fig.pos = 'h', fig.cap = "Plot of each microcosms surface cover over time for Hoare basin soils.  Counts begin on day 28 and end on day 63"}
plot(xyplot.h)
```
  
#### Plot growth by plate - Fryxell  

```{r echo = FALSE, include = FALSE}
#Fryxell
#xyplot for each plate
xyplot.f <- xyplot(primary_cover ~ days.f|plate_number, 
       data = df.f.1,
       groups = treatment, 
       xlab="Days since start of experiment",
       ylab="Surface Cover (%)",
       #main="Surface Cover by Plate- Bonney Basin Soils",
       type=c('g','p','l'),
       pch = 16, cex = 0.5,
       index.cond = function(x,y)max(y))

```
  
```{r echo = FALSE, fig.pos = 'h', fig.cap = "Plot of each microcosms surface cover over time for Fryxell basin soils.  Counts begin on day 28 and end on day 63"}
plot(xyplot.f)
```