---
title: "Nutrient Limitation in the MDV of Antarctica"
subtitle: "  \n Bonney Basin Surface Cover Over Time"
author: "Adam J. Solon"
date: "`r Sys.Date()`"
#output: html_document
output: 
  pdf_document:
    #includes:  
     # in_header: my_header.txt
    toc: TRUE
    fig_width: 8
    fig_height: 6
    fig_caption: TRUE
#fontsize: 12pt
#editor_options: 
#  chunk_output_type: console
---

# Script Summary  
This report plots surface growth of primary producers from Bonney Basin soils (interior basin) and uses statistical models to determine if there were differences in surface cover among treatments.  There are four treatment groups from each basin: CTRL (controls with H2O only), N (nitrogen + H2O only), P (phosphorus + H2O only), and NP (N and P combined + H2O). 

### Steps of this pipeline:  
1.  Create and organize directories
2.  Load R packages
4.  Input files
5.  Format Files
6.  Plot surface cover time series
7.  Plot maximum surface cover
8.  Statistics

```{r echo = FALSE, include = FALSE}
# Change identifiers to your system and file naming. 
user <- "C:/Users/adams" # computer user name
folder <- "/Desktop/Projects"
project <- "/MDV_Soils_nutrient_limitation/manuscript1"
analyses <- "/analyses"
data  <- "/point_intercept"
analysis2 <- "/SC_surface_cover"
analysis3 <- "/RR_response_ratio"
ecosystem <- "lab microcosms - Taylor Valley soils" # Define the environment for use in report language.
project.fp <- paste0(user, folder, project)
```

# Pipeline

### Set paths and create directories 

```{r set paths for project, echo = FALSE, include = FALSE}
# First define the project and project folders. 

# Create project directory
###################################################
project.fp <- paste0(user, folder, project)
if (!dir.exists(project.fp)) dir.create(project.fp)

# Create sub-directory for all analyses for project 
###################################################
analyses.fp <- paste0(project.fp, analyses)
if (!dir.exists(analyses.fp)) dir.create(analyses.fp)

# Create sub-directory for data type 
###################################################
data.fp <- paste0(analyses.fp, data)
if (!dir.exists(data.fp)) dir.create(data.fp)

# Create sub-directory for Surface Cover (SC) analysis
###################################################
analysis2.fp <- paste0(data.fp, analysis2)
if (!dir.exists(analysis2.fp)) dir.create(analysis2.fp)

# Create sub-directories for SC 
###################################################

# Create directory for 'before' pipeline inputs
input2.fp <- paste0(analysis2.fp, "/input")
if (!dir.exists(input2.fp)) dir.create(input2.fp)

# Create directory for 'within' pipeline R objects 
objects2.fp <- paste0(analysis2.fp, "/objects")
if (!dir.exists(objects2.fp)) dir.create(objects2.fp)

# Create directory for 'after' pipeline outputs 
output2.fp <- paste0(analysis2.fp, "/output")
if (!dir.exists(output2.fp)) dir.create(output2.fp)

# Check for the folders here as a sanity check. Should see "Input" and "Objects" if starting from scratch.
list.files(analysis2.fp) 

# Create further sub-directories by basin
###################################################

# Create directory for 'within' pipeline R objects 
objects2.b.fp <- paste0(objects2.fp, "/bonney")
if (!dir.exists(objects2.b.fp)) dir.create(objects2.b.fp)

# Create directory for 'after' pipeline outputs 
output2.b.fp <- paste0(output2.fp, "/bonney")
if (!dir.exists(output2.b.fp)) dir.create(output2.b.fp)

# Create sub-directory for Response Ratio (RR) analysis
###################################################
analysis3.fp <- paste0(data.fp, analysis3)
if (!dir.exists(analysis3.fp)) dir.create(analysis3.fp)

# Create sub-directory for RR 
###################################################

# Create directory for 'before' pipeline inputs
input3.fp <- paste0(analysis3.fp, "/input")
if (!dir.exists(input3.fp)) dir.create(input3.fp)

```

### Session Information  

```{r Install and load packages, echo = FALSE, include = FALSE}

# install.packages("tidyverse")
# install.packages("knitr")
# install.packages("kableExtra")
# install.packages("lattice")
# install.packages("betareg")
# install.packages("lmtest")
# install.packages("AICcmodavg")
# install.packages("boot")
# install.packages("emmeans")

library(tidyverse); packageVersion("tidyverse")
library(knitr); packageVersion("knitr")
library(kableExtra); packageVersion("kableExtra")
library(betareg); packageVersion("betareg")
library(lmtest); packageVersion("lmtest")
library(AICcmodavg); packageVersion("AICcmodavg")
library(boot); packageVersion("boot")
library(emmeans); packageVersion("emmeans")

knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(include = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

```  
  
* r version: `r getRversion()`
* RStudio version: `r rstudioapi::versionInfo()$version`
* r packages:  
  tidyverse, `r packageVersion("tidyverse")` 
  knitr, `r packageVersion("knitr")`  
  kableExtra, `r packageVersion("KableExtra")`  
  betareg, `r packageVersion("betareg")`  
  lmtest, `r packageVersion("lmtest")`  
  AICcmodavg, `r packageVersion("lmtest")`  
  boot, `r packageVersion("lmtest")`  
  emmeans, `r packageVersion("emmeans")`  
  
### Input file 
1. Count data from Bonney soil microcosms, filename 'Bonney_counts_percent.txt' an output from script 'MCosms_II_count_processing.Rmd'

```{r input data files, echo = FALSE, include = FALSE}
# input data files
# Count data from point-intercept method (microscopy)
counts.fp <- paste0(input2.fp, "/Bonney_counts_percent.txt") 

#input file
df <- read.table(counts.fp,  header = T, sep = "\t")

```

### Format file  

```{r echo = FALSE, include = FALSE}
# Reorder group factor levels
df$treatment <- factor(df$treatment,      
                           levels = c("CTRL", "N", "P", "NP"))

#remove over watered plate from analysis
df.1 <- df %>% filter(plate_number != 11)

```
  
### Run analyses

\newpage

# Plots  
  
## Surface Cover Over Time  
  
```{r echo = FALSE, include = FALSE}
#ggplot displaying growth by treatment
gc.plot.b <- ggplot(df.1, aes(x = days.b, y = primary_cover,
                   group = interaction(replicate, treatment),
                   colour = treatment)) +
  geom_point(alpha = 0.7, size = 3.5) +
  lims(y = c(0, 1)) +
  labs(x = "Days Since Treatment", 
       y = "Percent Cover (%)", 
       title = "Bonney") + 
  #subtitle = "Surface Cover per Week")+
  geom_line(linetype = 2, size = .65) +
  geom_jitter(height = 0, width = .5, show.legend = FALSE)+
  scale_x_continuous(expand = c(0.1, 1), limits = c(35, 70), breaks = c(35, 53, 70))+
  scale_y_continuous(expand = c(0, 0.025), limits = c(0, 100), breaks = c(0, 20, 40, 60, 80, 100)) +
  scale_color_manual(values = c("burlywood4",
                                "burlywood4",
                                "springgreen4",
                                "springgreen4")) +
  guides(color = "none", fill = "none") +
  theme_bw() + 
  theme(plot.title = element_text(size = rel(2.5)), 
        axis.title.y = element_blank(), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=18),
        axis.text.y = element_text(size=18),
        strip.text.x = element_text(size = 30),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank()) +#,
        #plot.margin = margin(t = 0,
         #                    b = 28,
          #                   l = 40,
           #                  r = 10)) +
  facet_grid(. ~ treatment) 

#save as R file
saveRDS(gc.plot.b, paste0(objects2.b.fp, "/BNM_surface_cover_plot.rds"))

```

``` {r echo = FALSE, fig.pos = 'h', fig.cap = "Plot of surface cover (%) by treatment for Bonney basin soils.  Counts begin on day 35 and end on day 70 and each dashed line represents a microcosm."}
#display graph
plot(gc.plot.b)
```
  
\newpage
  
## Maximum Surface Cover  
  
### Maximum Surface Cover by Treatment  
  
```{r echo = FALSE, include = FALSE}
#subset out only the largest surface cover achieved for each plate across the entire experiment
#### Controls
df.C <- subset(df.1, treatment == "CTRL") %>%
  group_by(plate_number) %>% 
  filter(primary_cover_2 == max(primary_cover_2, na.rm = TRUE))

# many equivalent max values for plate 15, remove superfluous values, 
df.C <- df.C[-c(5, 7), ]

#### Nitrogen
df.N <- subset(df.1, treatment == "N") %>%
  group_by(plate_number) %>% 
  filter(primary_cover_2 == max(primary_cover_2, na.rm = TRUE))

# many equivalent max values for plate 1, remove superfluous values
df.N <- df.N[-c(6), ]

#### Phosophorus
df.P <- subset(df.1, treatment == "P") %>%
  group_by(plate_number) %>% 
  filter(primary_cover_2 == max(primary_cover_2, na.rm = TRUE))

#### Nitrogen & Phosphorus
df.NP <- subset(df.1, treatment == "NP") %>%
  group_by(plate_number) %>% 
  filter(primary_cover_2 == max(primary_cover_2, na.rm = TRUE))

#join datafames by rows
df.2 <- rbind(df.C, df.N, df.P, df.NP)

#save as R file
saveRDS(df.2, paste0(objects2.b.fp, "/BNM_max_cover.Rds"))

#save as R file
saveRDS(df.2, paste0(input3.fp, "/BNM_max_cover.Rds"))

#save as .txt file 
write.table(df.2, file = paste0(output2.b.fp, "/BNM_max_cover.txt"), 
            sep = "\t", row.names = FALSE, col.names = TRUE)

```
  
```{r echo = FALSE, include = FALSE}
#input r object (max cover data frame) for plotting in graph
# df.2 <- readRDS(outputs.b.fp, "/BNM_max_cover.Rds")

#boxplots of averaged maximum cover by treatment
max.box.b <- df.2 %>% 
  ggplot(aes(x = treatment, y = primary_cover)) +
  geom_boxplot(color = "black",
               fill = c('burlywood4', "burlywood4", "springgreen4", "springgreen4"), 
               alpha = 0.3, 
               size = 1.0, 
               outlier.shape = NA) +
  geom_jitter(width = 0.2, 
              color = 'grey') +
  labs(x = "Treatment",
    y = "Percent Cover (%)"
  ) +
  scale_y_continuous(expand = c(0, 0.3), limits = c(0.0, 100.0), breaks=seq(0.0, 100.0, 10.0)) +
  scale_x_discrete(expand = c(0.04, 0.3)) +
  guides(color = "none") +
  theme_classic() +
  theme(plot.title = element_text(size = rel(2.5)),
        axis.title.y = element_text(size = 20), 
        axis.title.x = element_text(size = 20),
        axis.text.x = element_text(size = 18),
        axis.text.y = element_text(size = 18))

#ggplot of max cover per microcosm and grouped by treatment
max.plot.b <- ggplot() +
  geom_point(aes(x = replicate, y = primary_cover_2), data = df.2) +
  scale_y_continuous(expand = c(0, 0.025), limits = c(0, 1.0), breaks = c(0.0, .20, .40, .60, .80, 1.00)) +
  facet_grid(. ~ treatment, switch = "x") +
  theme(
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 11),
    axis.title.y = element_text(size = 11),
    axis.title.x = element_text(size = 11),
    panel.background = element_rect(fill = "white", colour = NA),
    panel.border = element_rect(fill = NA, colour = "grey50"),
    panel.grid.major = element_line(colour = "grey90", size = 0.2),
    panel.grid.minor = element_line(colour = "grey98", size = 0.5),
    strip.background = element_rect(fill = "grey80", colour = "grey50"),
    legend.text = element_text(size = 11),
    strip.text.x = element_text(size = 11)
  ) +
  ylab("Proportion cover") +
  xlab("Replicate no. in Treatment")
```
  
```{r echo = FALSE, fig.pos = 'h', fig.cap = "Plot of average maximum surface cover (%) by treatment for Bonney basin soils."}
#display
plot(max.box.b)
```
  
```{r echo=FALSE, include=FALSE}
#check and set working directory
setwd("C:/Users/adams/Desktop/NAM/manuscript1/analyses/point_intercept/SC_surface_cover/")
getwd()

#Save as jpg
jpeg(filename = "output/bonney/max_surface_cover_boxplot.jpeg", 
     width = 800, 
     height = 600, 
     units = "px", 
     pointsize = 12, 
     quality = 100, 
     bg = "white", 
     res = NA)
#
max.box.b
#
# Close device
dev.off()
```  
  
\newpage  
### Maximum Surface Cover for each Microcosm  
   
```{r echo = FALSE, fig.pos = 'h', fig.cap = "Plot of maximum surface cover for each microcosm grouped by treatment for Bonney basin soils."}
#display
plot(max.plot.b)
```
  
\newpage   
# Statistics  
  
## Does treatment explain surface cover growth?  
  
## Models  
  
### Beta regresion null model  
  
```{r echo = FALSE}
#### beta regression ####
#beta regression null model
b.null <- betareg(primary_cover_2 ~ 1, data = df.2)

summary(b.null)
```
  
### Beta regression model  
  
```{r echo = FALSE}
# explanatory variable = 'treatment'
b.reg <- betareg(primary_cover_2 ~ treatment, data = df.2)

#summary
summary(b.reg)

```
  
### Likelihood-ratio test  
  
```{r echo = FALSE}
#comparison of models using asymptotic likelihood ratio tests, used in comparing nested (generalized) linear models 
lrt <- lrtest(b.null, b.reg) ##likelihood-ratio test significant difference between null model and treatment model. 

#summary
print(lrt)

```

### Linear model  
  
```{r echo = FALSE}
#linear model
l.mod <- lm(primary_cover_2 ~ treatment, data = df.2)

summary(l.mod)

```
  
\newpage  
## Diagnostics  
### Plots  
  
```{r echo = FALSE, fig.pos = 'h', fig.cap = "Beta regression: fitted - raw residuals"}
#beta regression model- fitted vs. residuals
plot(fitted(b.reg),
     residuals(b.reg))
```
  
```{r echo = FALSE, fig.pos = 'h', fig.cap = "Linear model: fitted - raw residuals"}
#linear model- fitted vs. residuals
plot(fitted(l.mod),
     residuals(l.mod))
```
  
```{r echo = FALSE, fig.pos = 'h', fig.cap = "Beta regression: observations vs. pearson residuals"}
#observations vs. pearson residuals
plot(b.reg, which = 1, type = "pearson")
```
  
```{r echo = FALSE, fig.pos = 'h', fig.cap = "Beta regression: observations vs. deviance residuals"}
#observations vs. deviance residuals
plot(b.reg, which = 1, type = "deviance", sub.caption = "")
```
  
```{r echo = FALSE, fig.pos = 'h', fig.cap = "Beta regression: Cook's distance"}
#Cook's distance (pearson residuals)
plot(b.reg, which = 2)
```
  
``` {r echo = FALSE, fig.pos = 'h', fig.cap = "Linear Model: Cook's distance"}
#plot of cook's distance
plot(l.mod, which = 4)
```
  
```{r echo = FALSE, fig.pos = 'h', fig.cap = "Beta regression: Generalized leverage vs. predicted values"}
#Generalized leverage vs. predicted values
plot(b.reg, which = 3, type = "pearson")
```
 
```{r echo = FALSE, fig.pos = 'h', fig.cap = "Linear Model: Scale-Location"}
#scale location plot
plot(l.mod, which = 3)
```
  
```{r echo = FALSE, fig.pos = 'h', fig.cap = "Beta regression: pearson residuals vs. linear predictor"}
#pearson residuals vs. linear predictor
plot(b.reg, which = 4, type = "pearson")
```
  
```{r echo = FALSE, fig.pos = 'h', fig.cap = "Beta regression: half-normal plot (normal quantiles vs. deviance residuals)"}
#half-normal plot (normal quantiles) vs. deviance residuals
plot(b.reg, which = 5, type = "deviance", sub.caption = "")
```
 
```{r echo = FALSE, fig.pos = 'h', fig.cap = "Linear Model: Normal Q-Q plot (theoretical quantiles vs. standardized residuals)"}
#normal Q-Q plot
plot(l.mod, which = 2)
```
  
\newpage
  
## Model Selection  
  
### AIC  
  
```{r echo = FALSE, include = FALSE}
#model comparison with Akaike Information Criterion
aic <- AIC(b.reg, l.mod, b.null) 

#display
print(aic)

```
  
```{r echo = FALSE, include = FALSE}
#second order AICc (small sample size)
aicC.b <- AICc(b.reg, return.K = FALSE, second.ord = TRUE) 

#display
print(aicC.b)

#parameters in model
aicC.b.p <- AICc(b.reg, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.b.p)
```
  
```{r echo = FALSE, include = FALSE}
#second order AICc (small sample size)
aicC.l <- AICc(l.mod, return.K = FALSE, second.ord = TRUE)

#display
print(aicC.l)

#parameters in model
aicC.l.p <- AICc(l.mod, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.l.p)
```
  
```{r echo = FALSE, include = FALSE}
#second order AICc (small sample size)
aicC.n <- AICc(b.null, return.K = FALSE, second.ord = TRUE)

#display
print(aicC.n)

#parameters in model
aicC.n.p <- AICc(b.null, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.n.p)
```
  
```{r echo = FALSE, include = FALSE}
#create data frame w/ AICc results
aicC <- data.frame (Model  = c("beta.regression", "linear.model", "beta.null"),
                  AICc = c(aicC.b, aicC.l, aicC.n),
                  parameters = c(aicC.b.p, aicC.l.p, aicC.n.p)
                  )
```
  
```{r echo = FALSE, include = FALSE}
#subest columns for use in table for pdf
aic <- aic %>% rownames_to_column("Model")

#reorder columns
aic <- aic %>% relocate(AIC, .before = df)
```
  
```{r kable 1}
#create table of AIC values
knitr::kable(aic, booktabs = T, digits = 2, align = "ccc", caption = 'Model Comparison with Akaike Information Criterion (AIC)') %>%
  kable_styling(latex_options = c("striped", "hold_position"), font_size = 14) %>%
  row_spec(0, bold = T)

```
  
\newpage  
  
### AICc  
  
```{r kable 2}
#create table of AICc values
knitr::kable(aicC, booktabs = T, digits = 2, align = "ccc", caption = 'Model Comparison with  second-order Akaike Information Criterion (AICc), a correction for small sample size') %>%
  kable_styling(latex_options = c("striped", "hold_position"), font_size = 14) %>%
  row_spec(0, bold = T)

```
  
###### AIC analysis suggests improved model fit w/ beta regression relative to a basic linear model (i.e. ANOVA) or a null model.   
  
\newpage  
  
### Beta regression model 
###### Current beta regression model has fixed precision parameter.  Build model with a precision parameter modeled from explanatory variables (i.e., a variable dispersion beta regression) and compare models.
  
```{r echo = FALSE}
#repeat previous anlyses using a beta regression model that also allows ϕ to vary
b.reg.pp <- betareg(primary_cover_2 ~ treatment | treatment, data = df.2)
#
summary(b.reg.pp)
```
 
  
```{r echo = FALSE}
#Residual plots of b.reg and b.reg.pp
par(mfrow = c(2, 1), oma = c(0, 0, 0, 0), mar = c(4, 4, 0.2, 0.2))
plot(residuals(b.reg) ~ fitted(b.reg))
plot(residuals(b.reg.pp) ~ fitted(b.reg.pp))
``` 
    
###### Likelihood ratio tests comparing null, beta regression, and variable dispersion beta regression non-corrected, bias correction, and bias reduction
   
```{r echo = FALSE}
#estimators of the beta regression model may be biased, particularly when the sample size is small (n = 16 here when using patch-level observations)
#bias correction
b.null.bc <- betareg(primary_cover_2 ~ 1, data = df.2, type = "BC")
b.reg.bc <- betareg(primary_cover_2 ~ treatment, data = df.2, type = "BC")
b.reg.pp.bc <- update(b.reg.bc, . ~ . | treatment, type = "BC")

#bias reduction
b.null.br <- betareg(primary_cover_2 ~ 1, data = df.2, type = "BR")
b.reg.br <- betareg(primary_cover_2 ~ treatment, data = df.2, type = "BR")
b.reg.pp.br <- update(b.reg.br, . ~ . | treatment, type = "BR")

#likelihood ratio tests
lrtest(b.null, b.reg, b.reg.pp)
lrtest(b.null.bc, b.reg.bc, b.reg.pp.bc)
lrtest(b.null.br, b.reg.br, b.reg.pp.br)

```
  
\newpage  
### AICc  - Model Comparison
  
```{r echo = FALSE, include = FALSE}
#second order AICc (small sample size)
aicC.b <- AICc(b.reg, return.K = FALSE, second.ord = TRUE) 

#display
print(aicC.b)

#parameters in model
aicC.b.p <- AICc(b.reg, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.b.p)
```
  
```{r echo = FALSE, include = FALSE}
#second order AICc (small sample size)
aicC.n <- AICc(b.null, return.K = FALSE, second.ord = TRUE)

#display
print(aicC.n)

#parameters in model
aicC.n.p <- AICc(b.null, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.n.p)
```
  
```{r echo = FALSE, include = FALSE}
#second order AICc (small sample size)
aicC.b.pp <- AICc(b.reg.pp, return.K = FALSE, second.ord = TRUE) 

#display
print(aicC.b.pp)

#parameters in model
aicC.b.pp.p <- AICc(b.reg.pp, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.b.pp.p)
```
  
```{r echo = FALSE, include = FALSE}
#create data frame w/ AICc results
aicC.b.df <- data.frame (Model  = c("beta.regression", "var.disp.beta.reg", "beta.null"),
                  AICc = c(aicC.b, aicC.b.pp, aicC.n),
                  parameters = c(aicC.b.p, aicC.b.pp.p, aicC.n.p)
                  )
```
  
```{r kable 3}
#create table of AICc values
knitr::kable(aicC.b.df, booktabs = T, digits = 2, align = "ccc", caption = 'Model Comparison with  second-order Akaike Information Criterion (AICc), a correction for small sample size') %>%
  kable_styling(latex_options = c("striped", "hold_position"), font_size = 14) %>%
  row_spec(0, bold = T)

```
  
\newpage  
### AICc - Models with Bias Correction  
  
```{r echo = FALSE, include = FALSE}
#second order AICc (small sample size)
aicC.b.bc <- AICc(b.reg.bc, return.K = FALSE, second.ord = TRUE) 

#display
print(aicC.b.bc)

#parameters in model
aicC.b.bc.p <- AICc(b.reg.bc, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.b.bc.p)
```
  
```{r echo = FALSE, include = FALSE}
#second order AICc (small sample size)
aicC.n.bc <- AICc(b.null.bc, return.K = FALSE, second.ord = TRUE)

#display
print(aicC.n.bc)

#parameters in model
aicC.n.bc.p <- AICc(b.null.bc, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.n.bc.p)
```
  
```{r echo = FALSE, include = FALSE}
#second order AICc (small sample size)
aicC.b.pp.bc <- AICc(b.reg.pp.bc, return.K = FALSE, second.ord = TRUE) 

#display
print(aicC.b.pp.bc)

#parameters in model
aicC.b.pp.bc.p <- AICc(b.reg.pp.bc, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.b.pp.bc.p)
```
  
```{r echo = FALSE, include = FALSE}
#create data frame w/ AICc results
aicC.b.bc.df <- data.frame (Model  = c("beta.regression", "var.disp.beta.reg", "beta.null"),
                  AICc = c(aicC.b.bc, aicC.b.pp.bc, aicC.n.bc),
                  parameters = c(aicC.b.bc.p, aicC.b.pp.bc.p, aicC.n.bc.p)
                  )
```
  
```{r kable 4}
#create table of AICc values
knitr::kable(aicC.b.bc.df, booktabs = T, digits = 2, align = "ccc", caption = 'Bias Correction Model Comparison with  second-order Akaike Information Criterion (AICc), a correction for small sample size') %>%
  kable_styling(latex_options = c("striped", "hold_position"), font_size = 14) %>%
  row_spec(0, bold = T)

```
  
\newpage  
### AICc - Models with Bias Reduction  
  
```{r echo = FALSE, include = FALSE}
#second order AICc (small sample size)
aicC.b.br <- AICc(b.reg.br, return.K = FALSE, second.ord = TRUE) 

#display
print(aicC.b.br)

#parameters in model
aicC.b.br.p <- AICc(b.reg.br, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.b.br.p)
```
  
```{r echo = FALSE, include = FALSE}
#second order AICc (small sample size)
aicC.n.br <- AICc(b.null.br, return.K = FALSE, second.ord = TRUE)

#display
print(aicC.n.br)

#parameters in model
aicC.n.br.p <- AICc(b.null.br, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.n.br.p)
```
  
```{r echo = FALSE, include = FALSE}
#second order AICc (small sample size)
aicC.b.pp.br <- AICc(b.reg.pp.br, return.K = FALSE, second.ord = TRUE) 

#display
print(aicC.b.pp.br)

#parameters in model
aicC.b.pp.br.p <- AICc(b.reg.pp.br, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.b.pp.br.p)
```
  
```{r echo = FALSE}
#create data frame w/ AICc results
aicC.b.br.df <- data.frame (Model  = c("beta.regression", "var.disp.beta.reg", "beta.null"),
                  AICc = c(aicC.b.br, aicC.b.pp.br, aicC.n.br),
                  parameters = c(aicC.b.br.p, aicC.b.pp.br.p, aicC.n.br.p)
                  )
```
  
```{r kable 5}
#create table of AICc values
knitr::kable(aicC.b.br.df, booktabs = T, digits = 2, align = "ccc", caption = 'Bias Reduction Model Comparison with  second-order Akaike Information Criterion (AICc), a correction for small sample size') %>%
  kable_styling(latex_options = c("striped", "hold_position"), font_size = 14) %>%
  row_spec(0, bold = T)

```
  
\newpage  
  
###### Model Summaries of beta regression with bias correction or reduction, or basic model
```{r echo = FALSE}
#summary of model statistics
summary(b.reg.bc)
summary(b.reg.br)
summary(b.reg)

```
  
Final model selection using the principle of parsimony is the basic beta regression. There was increased surface cover of P and NP treatments as compared to CTRL.  Follow with a Post Hoc test for pairwise comparison of fitted model using estimated marginal means.  
  
```{r echo=FALSE}
#save as .txt file
capture.output(summary(b.reg), file = paste0(output2.b.fp, '/BNM_beta_regression_summary.txt'))
```

```{r echo = FALSE}
#post-hoc tests to identify which pairwise contrasts of treatments levels are significant
p.hoc <- test(pairs(emmeans(b.reg, ~ treatment, mode = "link")))

#display
print(p.hoc)

#save as .csv file
write.csv(p.hoc, file = paste0(output2.b.fp, '/BNM_betaReg_pairwise_test.csv'))
```
  
\newpage  
  
## Is there an interaction effect between factors (i.e., factor 1: nitrogen, and factor 2; phosphorus)  
  
## Models  
```{r echo = FALSE}
#model with explanatory variables = factor 1 (N), factor 2 (P) for interaction effects
b.reg.f <- betareg(primary_cover_2 ~ N * P, data = df.2)

#summary
summary(b.reg.f)

```
  
```{r echo = FALSE}
#model with explanatory variables = factor 1 (N), factor 2 (P) for interaction effects
b.reg.f.pp <- betareg(primary_cover_2 ~ N * P | N * P, data = df.2)

#summary
summary(b.reg.f.pp)

```
  
```{r echo = FALSE}
#Residual plots of b.reg and b.reg.pp
par(mfrow = c(2, 1), oma = c(0, 0, 0, 0), mar = c(4, 4, 0.2, 0.2))
plot(residuals(b.reg.f) ~ fitted(b.reg.f))
plot(residuals(b.reg.f.pp) ~ fitted(b.reg.f.pp))
```
 
```{r echo = FALSE, include = FALSE}
#second order AICc (small sample size)
aicC.b.f <- AICc(b.reg.f, return.K = FALSE, second.ord = TRUE) 

#display
print(aicC.b.f)

#parameters in model
aicC.b.f.p <- AICc(b.reg.f, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.b.f.p)
```
  
```{r echo = FALSE, include = FALSE}
#second order AICc (small sample size)
aicC.b.pp.f <- AICc(b.reg.f.pp, return.K = FALSE, second.ord = TRUE) 

#display
print(aicC.b.pp.f)

#parameters in model
aicC.b.pp.p.f <- AICc(b.reg.f.pp, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.b.pp.p.f)
```
  
```{r echo = FALSE, include = FALSE}
#create data frame w/ AICc results
aicC.b.f.df <- data.frame (Model  = c("beta.regression", "var.disp.beta.reg"),
                  AICc = c(aicC.b.f, aicC.b.pp.f),
                  parameters = c(aicC.b.f.p, aicC.b.pp.p.f)
                  )
```
  
```{r kable 6}
#create table of AICc values
knitr::kable(aicC.b.f.df, booktabs = T, digits = 2, align = "cc", caption = 'Model Comparison with  second-order Akaike Information Criterion (AICc), a correction for small sample size') %>%
  kable_styling(latex_options = c("striped", "hold_position"), font_size = 14) %>%
  row_spec(0, bold = T)

```

Model selection using the principle of parsimony suggests there was not an interaction effect between the N and P treatments. Follow with a Post Hoc test for pairwise comparison of fitted model.  

```{r echo=FALSE}
#save as .txt file
capture.output(summary(b.reg.f), file = paste0(output2.b.fp, '/BNM_beta_regression_interaction_summary.txt'))
```
  
```{r echo = FALSE}
#post-hoc test of model to identify which pairwise contrasts of treatments levels are significant
p.hoc.f <- test(pairs(emmeans(b.reg.f, ~ N * P, mode = "link")))

#display
print(p.hoc.f)

#save as .csv file
write.csv(p.hoc.f, file = paste0(output2.b.fp, '/BNM_betaReg_interaction_pairwise_test.csv'))
```
  
```{r echo = FALSE, fig.pos = 'h', fig.cap = "Interaction plot between N and P factors for Bonney basin soils."}
#display interaction
emmip(b.reg.f, P ~ N)
```

