---
title: "Nutrient Limitation in the MDV of Antarctica"
subtitle: "  \n Taylor Valley Surface Cover"
author: "Adam J. Solon"
date: "`r Sys.Date()`"
#output: html_document
output: 
  pdf_document:
    #includes:  
     # in_header: my_header.txt
    toc: TRUE
    fig_width: 8
    fig_height: 6
    fig_caption: TRUE
#fontsize: 12pt
#editor_options: 
#  chunk_output_type: console
---

# Script Summary  
This pipeline creates a figure for a manuscript on the growth of primary producers from nutrient enrichments in microcosms of Taylor Valley soils collected across three enclosed lake basins, Bonney (interior), Hoare (intermediate), and Fryxell (coastal), compared to the growth of the controls (i.e. no nutrients added). There are four treatment groups from each basin: CTRL (controls with H2O only), N (nitrogen + H2O only), P (phosphorus + H2O only), and NP (N and P combined + H2O).  This figure depicts the time series of surface growth as percent surface cover.

### Steps of this pipeline:  
1.  Create and organize directories
2.  Load R packages
3.  Input files
4.  Format File
5.  Run analyses
6.  Statistics
7.  Interaction Plot

```{r echo = FALSE, include = FALSE}
# Change identifiers to your system and file naming. 
user <- "C:/Users/adams" # computer user name
folder <- "/Desktop/Projects"
project <- "/MDV_Soils_nutrient_limitation/manuscript1"
analyses <- "/analyses"
data  <- "/point_intercept"
analysis2 <- "/SC_surface_cover"
ecosystem <- "lab microcosms - Taylor Valley soils" # Define the environment for use in report language.
project.fp <- paste0(user, folder, project)
```

# Pipeline

### Set paths and create directories 

```{r set paths for project, echo = FALSE, include = FALSE}
# First define the project and project folders. 

# Create project directory
###################################################
project.fp <- paste0(user, folder, project)
if (!dir.exists(project.fp)) dir.create(project.fp)

# Create sub-directory for all analyses for project 
###################################################
analyses.fp <- paste0(project.fp, analyses)
if (!dir.exists(analyses.fp)) dir.create(analyses.fp)

# Create sub-directory for data type 
###################################################
data.fp <- paste0(analyses.fp, data)
if (!dir.exists(data.fp)) dir.create(data.fp)

# Create sub-directory for Response Ratio (RR) analysis
###################################################
analysis2.fp <- paste0(data.fp, analysis2)
if (!dir.exists(analysis2.fp)) dir.create(analysis2.fp)

# Create sub-directories for RR 
###################################################

# Create directory for 'before' pipeline inputs
input2.fp <- paste0(analysis2.fp, "/input")
if (!dir.exists(input2.fp)) dir.create(input2.fp)

# Create directory for 'within' pipeline R objects 
objects2.fp <- paste0(analysis2.fp, "/objects")
if (!dir.exists(objects2.fp)) dir.create(objects2.fp)

# Create directory for 'after' pipeline outputs 
output2.fp <- paste0(analysis2.fp, "/output")
if (!dir.exists(output2.fp)) dir.create(output2.fp)

# Create directory for 'figures' for manuscript 
figures2.fp <- paste0(analysis2.fp, "/figures")
if (!dir.exists(figures2.fp)) dir.create(figures2.fp)

# create directories for each basin in objects directory
#####################################################
# Create directory for bonney 
objects2.b.fp <- paste0(objects2.fp, "/bonney")
if (!dir.exists(objects2.b.fp)) dir.create(objects2.b.fp)

# Create directory for hoare
objects2.h.fp <- paste0(objects2.fp, "/hoare")
if (!dir.exists(objects2.h.fp)) dir.create(objects2.h.fp)

# Create directory for fryxell 
objects2.f.fp <- paste0(objects2.fp, "/fryxell")
if (!dir.exists(objects2.f.fp)) dir.create(objects2.f.fp)

# Check for the folders here as a sanity check. Should see "Input" and "Objects" if starting from scratch.
list.files(analysis2.fp) 

```

### Session Information  

```{r Install and load packages, echo = FALSE, include = FALSE}

# install.packages("tidyverse")
# install.packages("knitr")
# install.packages("ggpubr")

library(tidyverse); packageVersion("tidyverse")
library(knitr); packageVersion("knitr")
library(ggpubr); packageVersion("ggpubr")

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

```  
  
* r version: `r getRversion()`
* RStudio version: `r rstudioapi::versionInfo()$version`
* r packages:  
  tidyverse, `r packageVersion("tidyverse")` 
  knitr, `r packageVersion("knitr")`  
  kableExtra, `r packageVersion("KableExtra")`  
  
### Input files 
1. R objects of figures for each basin, filenames 'BNM_surface_cover_plot.Rds' , 'HRM_surface_cover_plot.Rds', and 'FXM_surface_cover_plot.Rds' outputs from scripts 'MCosms_II_surface_cover_bonney.Rmd', 'MCosms_II_surface_cover_hoare.Rmd', and 'MCosms_II_surface_cover_fryxell.Rmd'  

```{r input data files, echo = FALSE, include = FALSE}
# input data files
# figures of respone ratios
figure2.b.fp <- paste0(objects2.b.fp, "/BNM_surface_cover_plot.Rds") 
figure2.h.fp <- paste0(objects2.h.fp, "/HRM_surface_cover_plot.Rds") 
figure2.f.fp <- paste0(objects2.f.fp, "/FXM_surface_cover_plot.Rds") 


#input files
fg2.b <- readRDS(figure2.b.fp)
fg2.h <- readRDS(figure2.h.fp)
fg2.f <- readRDS(figure2.f.fp)

```
  
### Arrange figures into one combined figure for manuscript  
  
\newpage
```{r combine figures, echo=FALSE, include=FALSE}
#### arrange all lnRR interaction plots from each basin together in one figure
fg2.all <- ggarrange(fg2.b, fg2.h, fg2.f,
          #labels = c("A", "B", "C"),
          nrow = 3,
          align = "hv")

```
  
```{r time series plot, echo=FALSE, fig.height=12, fig.width=8, fig.pos = 'h', fig.cap = "Surface Growth as percent cover of soils from three enclosed lakes basins within Taylor Valley enriched with N and/or P, or no enrichment (i.e., controls). Each dashed line represents a single microcosm (i.e., replicate)."}
#plot
fg2.all

```
  
```{r save combined figure as jpeg, echo=FALSE, include=FALSE}
#check and set working directory
setwd("C:/Users/adams/Desktop/projects/MDV_Soils_nutrient_limitation/manuscript1/analyses/point_intercept/SC_surface_cover/")
getwd()

#Save as jpg
jpeg(filename = "figures/Taylor_Valley_surface_cover_time_series_plot.jpeg", 
     width = 800, 
     height = 1200, 
     units = "px", 
     pointsize = 11, 
     quality = 100, 
     bg = "white", 
     res = NA)
#
fg2.all
#
# Close device
dev.off()
```
