---
title: "Nutrient Limitation in the MDV of Antarctica"
subtitle: "  \n Hoare Basin Response Ratios"
author: "Adam J. Solon"
date: "`r Sys.Date()`"
#output: html_document
output: 
  pdf_document:
    #includes:  
     # in_header: my_header.txt
    toc: TRUE
    fig_width: 8
    fig_height: 6
    fig_caption: TRUE
#fontsize: 12pt
#editor_options: 
#  chunk_output_type: console
---

# Script Summary  
This report quantifies the response of the growth of primary producers from nutrient enrichments in Hoare Basin soils compared to the growth of the controls (i.e. no nutrients added). There are four treatment groups from each basin: CTRL (controls with H2O only), N (nitrogen + H2O only), P (phosphorus + H2O only), and NP (N and P combined + H2O). Response ratios provide a quantitative effect size metric that allows for comparisons within and across basins.

### Steps of this pipeline:  
1.  Create and organize directories
2.  Load R packages
3.  Input files
4.  Format File
5.  Run analyses
6.  Statistics
7.  Plot  
  
```{r echo = FALSE, include = FALSE}
# Change identifiers to your system and file naming. 
user <- "C:/Users/adams" # computer user name
folder <- "/Desktop/Projects"
project <- "/MDV_Soils_nutrient_limitation/manuscript1"
analyses <- "/analyses"
data  <- "/point_intercept"
analysis3 <- "/RR_response_ratio"
ecosystem <- "lab microcosms - Taylor Valley soils" # Define the environment for use in report language.
project.fp <- paste0(user, folder, project)
```

# Pipeline

### Set paths and create directories 

```{r set paths for project, echo = FALSE, include = FALSE}
# First define the project and project folders. 

# Create project directory
###################################################
project.fp <- paste0(user, folder, project)
if (!dir.exists(project.fp)) dir.create(project.fp)

# Create sub-directory for all analyses for project 
###################################################
analyses.fp <- paste0(project.fp, analyses)
if (!dir.exists(analyses.fp)) dir.create(analyses.fp)

# Create sub-directory for data type 
###################################################
data.fp <- paste0(analyses.fp, data)
if (!dir.exists(data.fp)) dir.create(data.fp)

# Create sub-directory for Response Ratio (RR) analysis
###################################################
analysis3.fp <- paste0(data.fp, analysis3)
if (!dir.exists(analysis3.fp)) dir.create(analysis3.fp)

# Create sub-directories for RR 
###################################################

# Create directory for 'before' pipeline inputs
input3.fp <- paste0(analysis3.fp, "/input")
if (!dir.exists(input3.fp)) dir.create(input3.fp)

# Create directory for 'within' pipeline R objects 
objects3.fp <- paste0(analysis3.fp, "/objects")
if (!dir.exists(objects3.fp)) dir.create(objects3.fp)

# Create directory for 'after' pipeline outputs 
output3.fp <- paste0(analysis3.fp, "/output")
if (!dir.exists(output3.fp)) dir.create(output3.fp)

# Check for the folders here as a sanity check. Should see "Input" and "Objects" if starting from scratch.
list.files(analysis3.fp) 

# Create further sub-directories by basin
###################################################

# Create directory for 'within' pipeline R objects 
objects3.h.fp <- paste0(objects3.fp, "/hoare")
if (!dir.exists(objects3.h.fp)) dir.create(objects3.h.fp)

# Create directory for 'after' pipeline outputs 
output3.h.fp <- paste0(output3.fp, "/hoare")
if (!dir.exists(output3.h.fp)) dir.create(output3.h.fp)

```

### Session Information  

```{r Install and load packages, echo = FALSE, include = FALSE}

# install.packages("tidyverse")
# install.packages("knitr")
# install.packages("kableExtra")

library(tidyverse); packageVersion("tidyverse")
library(knitr); packageVersion("knitr")
library(kableExtra); packageVersion("kableExtra")
library(emmeans); packageVersion("emmeans")

knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(include = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

```  
  
* r version: `r getRversion()`
* RStudio version: `r rstudioapi::versionInfo()$version`
* r packages:  
  tidyverse, `r packageVersion("tidyverse")` 
  knitr, `r packageVersion("knitr")`  
  kableExtra, `r packageVersion("KableExtra")`  
  emmeans, `r packageVersion("emmeans")`
   
  
### Input file 
1. Maxmimum surface growth data from Hoare soil microcosms, filename '/HRM_max_cover.Rds' an output from script 'MCosms_II_surface_cover_hoare.Rmd'  

```{r input data files, echo = FALSE, include = FALSE}
# input data files
# Count data from point-intercept method (microscopy)
counts.fp <- paste0(input3.fp, "/HRM_max_cover.Rds") 

#input file
df <- readRDS(counts.fp)

```
  
### Format file  

```{r echo = FALSE, include = FALSE}
#reclassify plate_number and treatment as factors
df <- df %>% 
  mutate_at(vars(plate_number, N, P, treatment), factor)

# Reorder group factor levels
df$treatment <- factor(df$treatment,      
                           levels = c("CTRL", "N", "P", "NP"))

#subset controls
df.C <- subset(df, treatment == "CTRL") %>%
  dplyr::select(plate_number, treatment, primary_cover)

#subset N
df.N <- subset(df, treatment == "N") %>%
  dplyr::select(plate_number, treatment, primary_cover)

#subset P
df.P <- subset(df, treatment == "P") %>%
  dplyr::select(plate_number, treatment, primary_cover)

#subset NP
df.NP <- subset(df, treatment == "NP") %>%
  dplyr::select(plate_number, treatment, primary_cover)

```
  
### Run analyses  
  
# Statistics  
  
## Summary
#Control  
```{r echo = FALSE, include = FALSE}
#transform response variable with natural log
df.C$log <- log(df.C$primary_cover)  #ln of response variable
#address zeroes in dataset
df.C['log'][df.C['log'] == -Inf] <- 0

#compute the mean, standard deviation, sample size, standard error, and upper and lower 95% confidence intervals
C.mean <- mean(df.C$log)  
C.sd <- sd(df.C$log)  
C.size <- length(df.C$log) #number of replicates in treatment
C.se <- C.sd/sqrt(C.size) #standard error of the mean
C.cil <- C.mean - (C.se * 1.96) #lower bound of 95% confidence interval
C.ciu <- C.mean + (C.se * 1.96) #upper bound of 95% confidence interval

#join dataframe of vectors <mean, sd, se, ci>
dat.C <- data.frame(mean = C.mean, sd = C.sd, n= C.size, se = C.se, ci.low = C.cil, cu.up = C.ciu)
dat.C

#add column for treatment
dat.C$treatment <- "CTRL"

```
  
#Nitrogen
```{r echo = FALSE, include = FALSE}
#transform response variable with natural log
df.N$log <- log(df.N$primary_cover)  #ln of response variable

#compute the mean, standard deviation, sample size, standard error, and upper and lower 95% confidence intervals
N.mean <- mean(df.N$log)  
N.sd <- sd(df.N$log)  
N.size <- length(df.N$log) #number of replicates in treatment
N.se <- N.sd/sqrt(N.size) #standard error of the mean
N.cil <- N.mean - (N.se * 1.96) #lower bound of 95% confidence interval
N.ciu <- N.mean + (N.se * 1.96) #upper bound of 95% confidence interval

#join dataframe of vectors <mean, sd, se, ci>
dat.N <- data.frame(mean = N.mean, sd = N.sd, n= N.size, se = N.se, ci.low = N.cil, cu.up = N.ciu)
dat.N

#add column for treatment
dat.N$treatment <- "N"

```
  
#Phosphorus
```{r echo = FALSE, include = FALSE}
#transform response variable with natural log
df.P$log <- log(df.P$primary_cover)  #ln of response variable

#address zeroes in dataset
df.P['log'][df.P['log'] == -Inf] <- 0 #change -infinity value to zero

#compute the mean, standard deviation, sample size, standard error, and upper and lower 95% confidence intervals
P.mean <- mean(df.P$log)  
P.sd <- sd(df.P$log)  
P.size <- length(df.P$log) #number of replicates in treatment
P.se <- P.sd/sqrt(P.size) #standard error of the mean
P.cil <- P.mean - (P.se * 1.96) #lower bound of 95% confidence interval
P.ciu <- P.mean + (P.se * 1.96) #upper bound of 95% confidence interval

#join dataframe of vectors <mean, sd, se, ci>
dat.P <- data.frame(mean = P.mean, sd = P.sd, n= P.size, se = P.se, ci.low = P.cil, cu.up = P.ciu)
dat.P

#add column for treatment
dat.P$treatment <- "P"

```
  
#Nitrogen and Phosphorus
```{r echo = FALSE, include = FALSE}
#transform response variable with natural log
df.NP$log <- log(df.NP$primary_cover)  #ln of response variable

#compute the mean, standard deviation, sample size, standard error, and upper and lower 95% confidence intervals
NP.mean <- mean(df.NP$log)  
NP.sd <- sd(df.NP$log)  
NP.size <- length(df.NP$log) #number of replicates in treatment
NP.se <- NP.sd/sqrt(NP.size) #standard error of the mean
NP.cil <- NP.mean - (NP.se * 1.96) #lower bound of 95% confidence interval
NP.ciu <- NP.mean + (NP.se * 1.96) #upper bound of 95% confidence interval

#join dataframe of vectors <mean, sd, se, ci>
dat.NP <- data.frame(mean = NP.mean, sd = NP.sd, n= NP.size, se = NP.se, ci.low = NP.cil, cu.up = NP.ciu)
dat.NP

#add column for treatment
dat.NP$treatment <- "NP"

```
  
```{r echo = FALSE, include = FALSE}
#Combine data frames
dat.all <- rbind(dat.N, dat.P, dat.NP, dat.C)
dat.all

#save as csv file
write.csv(dat.all, paste0(output3.h.fp, '/HRM_summary_statistics.csv'))

```
  
```{r echo = FALSE, include = FALSE}
#Combine data frames
df.all <- rbind(df.C, df.N, df.P, df.NP)
df.all

#Save as R file
saveRDS(df.all, paste0(objects3.h.fp, "/HRM_df_all.rds"))

##save as .csv file
write.csv(df.all, file = paste0(output3.h.fp, '/HRM_summary_log_values.csv'))

```
  
The mean of the ln values for treatment C `r dat.C$mean`  
The mean of the ln values for treatment N `r dat.N$mean`  
The mean of the ln values for treatment P `r dat.P$mean`  
The mean of the ln values for treatment NP `r dat.NP$mean`   
  
Response Ratios:  
  
N - C = `r dat.N$mean - dat.C$mean`  
P - C = `r dat.P$mean - dat.C$mean`  
NP - C = `r dat.NP$mean - dat.C$mean`  
(N-C)+(P-C) = `r (dat.N$mean - dat.C$mean) + (dat.P$mean - dat.C$mean)`  
  
## Inferential  
Linear model  
```{r echo = FALSE}
#linear model
lm.LRR <- lm(log ~ treatment, data = df.all)

#model output
summary(lm.LRR)
```
  
Pairwise comparison (estimated marginal means)  
```{r echo = FALSE}
#post hoc pairwise comparison of fitted model
pairs(emmeans(lm.LRR, ~ treatment))

#save summary as .txt file
capture.output(summary(lm.LRR), file = paste0(output3.h.fp, "/HRM_lmLRR_summary.txt"))

#save pairwise test of fitted model (estimated marginal means) as .txt file
capture.output(pairs(emmeans(lm.LRR, ~ treatment)), file = paste0(output3.h.fp, "/HRM_lmLRR_pairwise.txt"))

```
  
```{r echo = FALSE, include = FALSE}
#check that model estimate is the mean lnN - mean lnC
dat.N$mean - dat.C$mean

#check that model estimate is the mean lnP - mean lnC
dat.P$mean - dat.C$mean

#check that model estimate is the mean lnNP - mean lnC
dat.NP$mean - dat.C$mean
```
  
Check estimates of model to ensure they are the same as the response ratio calculations (i.e., mean[ln X] - mean[ln C]).  
  
N - C = `r dat.N$mean - dat.C$mean`  
P - C = `r dat.P$mean - dat.C$mean`  
NP - C = `r dat.NP$mean - dat.C$mean`  
  
```{r echo = FALSE, include = FALSE}
#extract N treatment estimate (mean), standard error, and upper and lower confidence intervals from summary output
est <- coef(lm.LRR)[2]
se <- summary(lm.LRR)$coef[2, 2]
lcl <- (est - 1.96*se)
ucl <- (est + 1.96*se) 

#join vectors in in data frame
LRRn <- data.frame(est, se, lcl, ucl)
LRRn

#extract P treatment estimate (mean), standard error, and upper and lower confidence intervals from summary output
est <- coef(lm.LRR)[3]
se <- summary(lm.LRR)$coef[3, 2]
lcl <- (est - 1.96*se)
ucl <- (est + 1.96*se) 

##join vectors in in data frame
LRRp <- data.frame(est, se, lcl, ucl)
LRRp

#extract NP treatment estimate (mean), standard error, and upper and lower confidence intervals from summary output
est <- coef(lm.LRR)[4]
se <- summary(lm.LRR)$coef[4, 2]
lcl <- (est - 1.96*se)
ucl <- (est + 1.96*se) 

##join vectors in in data frame
LRRnp <- data.frame(est, se, lcl, ucl)
LRRnp

#add treatment column
LRRn$treatment <- c("N")
LRRp$treatment <- c("P")
LRRnp$treatment <- c("NP")

#Combine data frames
df.LRR <- rbind(LRRn, LRRp, LRRnp)
df.LRR

#save as .csv file
write.csv(df.LRR, file = paste0(output3.h.fp, '/HRM_dfLRR_summary_statistics.csv'))
```
  
```{r echo = FALSE, include = FALSE}
#interaction plot
#add new row for baseline- control
ctrl <- c(0.0, 0.0, 0.0, 0.0, "C")

#join vector with data frame
df.LRR <- rbind(ctrl, df.LRR)
df.LRR

#add row name for control row
row.names(df.LRR)[1] <- "TreatmentC"  

#create funtion to check or numeric 
is_all_numeric <- function(x) {
  !any(is.na(suppressWarnings(as.numeric(na.omit(x))))) & is.character(x)
}

#use function to change to numeric
df.LRR <- df.LRR %>% 
  mutate_if(is_all_numeric, as.numeric)

#round off LRR to two digits
df.LRR <- df.LRR %>% 
  mutate_if(is.numeric, round, digit = 3)

#add new columns by treatment factors
df.LRR$Nitrogen <- c("-N", "+N", "-N", "+N")
df.LRR$Phosphorus <- c("-P", "-P", "+P", "+P")

#reclassify treatment as factor with four levels
df.LRR$treatment <- as.factor(df.LRR$treatment)

#display
print(df.LRR)

#save as R file
saveRDS(df.LRR, paste0(objects3.h.fp, "/HRM_dfLRR.rds"))

```
  
\newpage
  
# Plot  

```{r echo = FALSE, include = FALSE}
#Restore dfLRR
#df.LRR <- readRDS(objects3.f.fp, "HRM_dfLRR.rds")

#plot
i.plot.h <- df.LRR %>% 
  ggplot(aes(x = Nitrogen, y=  est, 
             color = Phosphorus, 
             label = treatment)) +
  geom_line(size = 1.2, 
            aes(group = Phosphorus), 
            linetype = "twodash") +
  geom_point(size = 5.0, 
             alpha=0.7, 
             shape = 21, 
             fill = "black") +
  geom_errorbar(aes(ymin = lcl,
                    ymax = ucl),
                width = 0.03,
                size = 0.75,
                color = "grey") +
  geom_label(aes(x = 2.06, y = 1.85,
                 label = "N"),
             stat = "unique",
             fontface = "bold",
             size = 10, 
             color = "white", 
             fill = "burlywood4") +
  geom_label(aes(x = 0.95, y = 0.68,
                 label = "P"),
             stat = "unique",
             fontface = "bold", 
             size = 10, 
             color = "white", 
             fill = "springgreen4") +
  geom_label(aes(x = 2.08, y = 3.8,
                 label = "NP"),
             stat = "unique",
             fontface = "bold", 
             size = 10, 
             color = "white", 
             fill = "springgreen4") +
  geom_label(aes(x = 1, y = -0.2,
                 label = "C"),
             stat = "unique",
             fontface = "bold", 
             size = 10, 
             color = "white", 
             fill = "burlywood4") +
  labs(
    x = "Nitrogen",
    y = "lnRR "
  ) +
  scale_y_continuous(expand = c(0.01, 0.05), limits = c(-0.75, 4.50), breaks=seq(-0.75, 4.50, 0.75)) +
  scale_x_discrete(expand = c(0, 0.3)) +
  scale_color_manual(values = c("#999999",
                                "#333333")) +
  scale_fill_manual(values=c("burlywood4",
                             "burlywood4",
                             "springgreen4",
                             "springgreen4")) + 
  geom_hline(yintercept = 0, linetype="dashed", color = "red", size = 1.2) +
  geom_hline(yintercept = 2.49, linetype="dashed", color = "blue", size = 1.2) +
  guides(color = "none", fill = "none") +
  theme_bw() +
  theme(plot.title = element_text(size = 35), 
        axis.title.x = element_text(size = 30),
        axis.text.x = element_text(size = 25),
        axis.title.y = element_text(size = 30),
        axis.text.y = element_text(size = 25),
        axis.ticks.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank())

```
  
``` {r echo = FALSE, fig.pos = 'h', fig.cap = "Interaction plot between N and P of response ratios of nutrient enrichments vs. controls in Hoare basin soils. Red dashed line represents control and blue dashed line indicates an additive."}
#display graph
plot(i.plot.h)
```
  
```{r echo=FALSE, include=FALSE}
#check and set working directory
setwd("C:/Users/adams/Desktop/projects/MDV_Soils_nutrient_limitation/manuscript1/analyses/point_intercept/RR_response_ratio/")
getwd()

#Save as jpg
jpeg(filename = "output/hoare/HRM_response_ratio_interaction_plot.jpeg", 
     width = 800, 
     height = 600, 
     units = "px", 
     pointsize = 12, 
     quality = 100, 
     bg = "white", 
     res = NA)
#
i.plot.h
#
# Close device
dev.off()
```

```{r echo = FALSE, include = FALSE}
#plot for manuscript figure
i.plot.h.m <- df.LRR %>% 
  ggplot(aes(x = Nitrogen, y=  est, 
             color = Phosphorus, 
             label = treatment)) +
  geom_line(size = 1.2, 
            aes(group = Phosphorus), 
            linetype = "twodash") +
  geom_point(size = 5.0, 
             alpha = 0.7, 
             shape = 21, 
             fill = "black") +
  geom_errorbar(aes(ymin = lcl,
                    ymax = ucl),
                width = 0.03,
                size = 0.75,
                color = "grey") +
  geom_label(aes(fill = factor(treatment)), 
             colour = "white", 
             fontface = "bold", 
             size = 14, 
             nudge_y = 0.175,
             nudge_x = -0.19) +
  labs(
    title = "Hoare",
    #subtitle = paste0("Interaction between Nitrogen and Phosphorus Enrichments"),
    x = "Nitrogen",
    y = "lnRR "
  ) +
  scale_y_continuous(expand = c(0.01, 0.05), limits = c(-0.75, 4.50), breaks=seq(-0.75, 4.50, 0.75)) +
  scale_x_discrete(expand = c(0, 0.3)) +
  scale_color_manual(values = c("#999999",
                                "#333333")) +
  scale_fill_manual(values=c("burlywood4",
                             "springgreen4",
                             "springgreen4",
                             "burlywood4")) + 
  geom_hline(yintercept = 0, linetype="dashed", color = "red", size = 1.2) +
  geom_hline(yintercept = 2.49, linetype="dashed", color = "blue", size = 1.2) +
  guides(color = "none", fill = "none") +
  theme_bw() +
  theme(plot.title = element_text(size = 40), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 35),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.y = element_blank(),
        plot.margin = margin(l = 80, # Left margin
                             r = 5,
                             b = 5,  # bottom margin
                             t = 10)) 

#save as R file
saveRDS(i.plot.h.m, paste0(objects3.h.fp, "/HRM_lnRR_plot_figure3.rds"))
```
  
## Is there single limitation or co-limitation?  If co-limitation, what type (e.g. super-additive, additive, or sub-additive)?   
  
The results suggest co-limitation and of super-additive (or synergistic) type.  
  
RRn + RRp = `r (dat.N$mean - dat.C$mean) + (dat.P$mean - dat.C$mean)`  
  
RRnp = `r dat.NP$mean - dat.C$mean`  