---
title: "Nutrient Limitation in the MDV of Antarctica"
subtitle: "  \n Taylor Valley Response Ratios"
author: "Adam J. Solon"
date: "`r Sys.Date()`"
#output: html_document
output: 
  pdf_document:
    #includes:  
     # in_header: my_header.txt
    toc: TRUE
    fig_width: 8
    fig_height: 6
    fig_caption: TRUE
#fontsize: 12pt
#editor_options: 
#  chunk_output_type: console
---

# Script Summary  
This pipeline creates a figure for a manuscript on the growth of primary producers from nutrient enrichment in microcosms of Taylor Valley soils collected across three enclosed lake basins, Bonney (interior), Hoare (intermediate), and Fryxell (coastal), compared to the growth of the controls (i.e. no nutrients added). There are four treatment groups from each basin: CTRL (controls with H2O only), N (nitrogen + H2O only), P (phosphorus + H2O only), and NP (N and P combined + H2O).  This figure depicts the response ratios which compare the growth of enriched soils vs. control soils (i.e. no nutrient(s) added) and provide a quantitative effect size metric that allows for comparisons within and across basins.  
  
### Steps of this pipeline:  
1.  Create and organize directories
2.  Load R packages
3.  Input files
4.  Format File
5.  Run analyses
6.  Statistics
7.  Interaction Plot

```{r echo = FALSE, include = FALSE}
# Change identifiers to your system and file naming. 
user <- "C:/Users/adams" # computer user name
folder <- "/Desktop/Projects"
project <- "/MDV_Soils_nutrient_limitation/manuscript1"
analyses <- "/analyses"
data  <- "/point_intercept"
analysis3 <- "/RR_response_ratio"
ecosystem <- "lab microcosms - Taylor Valley soils" # Define the environment for use in report language.
project.fp <- paste0(user, folder, project)
```

# Pipeline

### Set paths and create directories 

```{r set paths for project, echo = FALSE, include = FALSE}
# First define the project and project folders. 

# Create project directory
###################################################
project.fp <- paste0(user, folder, project)
if (!dir.exists(project.fp)) dir.create(project.fp)

# Create sub-directory for all analyses for project 
###################################################
analyses.fp <- paste0(project.fp, analyses)
if (!dir.exists(analyses.fp)) dir.create(analyses.fp)

# Create sub-directory for data type 
###################################################
data.fp <- paste0(analyses.fp, data)
if (!dir.exists(data.fp)) dir.create(data.fp)

# Create sub-directory for Response Ratio (RR) analysis
###################################################
analysis3.fp <- paste0(data.fp, analysis3)
if (!dir.exists(analysis3.fp)) dir.create(analysis3.fp)

# Create sub-directories for RR 
###################################################

# Create directory for 'before' pipeline inputs
input3.fp <- paste0(analysis3.fp, "/input")
if (!dir.exists(input3.fp)) dir.create(input3.fp)

# Create directory for 'within' pipeline R objects 
objects3.fp <- paste0(analysis3.fp, "/objects")
if (!dir.exists(objects3.fp)) dir.create(objects3.fp)

# Create directory for 'after' pipeline outputs 
output3.fp <- paste0(analysis3.fp, "/output")
if (!dir.exists(output3.fp)) dir.create(output3.fp)

# Create directory for 'figures' for manuscript 
figures3.fp <- paste0(analysis3.fp, "/figures")
if (!dir.exists(figures3.fp)) dir.create(figures3.fp)

# create directories for each basin in objects directory
#####################################################
# Create directory for bonney 
objects3.b.fp <- paste0(objects3.fp, "/bonney")
if (!dir.exists(objects3.b.fp)) dir.create(objects3.b.fp)

# Create directory for hoare
objects3.h.fp <- paste0(objects3.fp, "/hoare")
if (!dir.exists(objects3.h.fp)) dir.create(objects3.h.fp)

# Create directory for fryxell 
objects3.f.fp <- paste0(objects3.fp, "/fryxell")
if (!dir.exists(objects3.f.fp)) dir.create(objects3.f.fp)

# Check for the folders here as a sanity check. Should see "Input" and "Objects" if starting from scratch.
list.files(analysis3.fp) 

```

### Session Information  

```{r Install and load packages, echo = FALSE, include = FALSE}

# install.packages("tidyverse")
# install.packages("knitr")
# install.packages("ggpubr")

library(tidyverse); packageVersion("tidyverse")
library(knitr); packageVersion("knitr")
library(ggpubr); packageVersion("ggpubr")

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

```  
  
* r version: `r getRversion()`
* RStudio version: `r rstudioapi::versionInfo()$version`
* r packages:  
  tidyverse, `r packageVersion("tidyverse")` 
  knitr, `r packageVersion("knitr")`  
  kableExtra, `r packageVersion("KableExtra")`  
  
### Input files 
1. R objects of figures for each basin, filenames 'BNM_lnRR_plot_figure3.Rds' , 'HRM_lnRR_plot_figure3.Rds', and 'FXM_lnRR_plot_figure3.Rds' outputs from scripts 'MCosms_II_response_ratio_bonney.Rmd', 'MCosms_II_response_ratio_hoare.Rmd', and 'MCosms_II_response_ratio_fryxell.Rmd'  

```{r input data files, echo = FALSE, include = FALSE}
# input data files
# figures of respone ratios
figure3.b.fp <- paste0(objects3.b.fp, "/BNM_lnRR_plot_figure3.Rds") 
figure3.h.fp <- paste0(objects3.h.fp, "/HRM_lnRR_plot_figure3.Rds") 
figure3.f.fp <- paste0(objects3.f.fp, "/FXM_lnRR_plot_figure3.Rds") 


#input files
fg3.b <- readRDS(figure3.b.fp)
fg3.h <- readRDS(figure3.h.fp)
fg3.f <- readRDS(figure3.f.fp)

```
  
### Arrange figures into one combined figure for manuscript  
```{r combine figures, echo=FALSE, include=FALSE}
#### arrange all lnRR interaction plots from each basin together in one figure
fg3.all <- ggarrange(fg3.b, fg3.h, fg3.f,
          #labels = c("A", "B", "C"),
          ncol = 3)

```
  
```{r respone ratio figure, echo=FALSE, fig.width = 12, fig.height = 6, fig.pos = 'h', fig.cap = "Response Ratio Interaction Plots of soils from three enclosed lakes basins within Taylor Valley enriched with N and/or P, or no enrichment (i.e., C - controls). Red dashed line represents control and blue dashed line indicates an additive."}
#plot
fg3.all

```
### Save combined figure as jpeg
```{r save combined figure as jpeg, echo=FALSE, include=FALSE}
#check and set working directory
setwd("C:/Users/adams/Desktop/projects/MDV_Soils_nutrient_limitation/manuscript1/analyses/point_intercept/RR_response_ratio/")
getwd()

#Save as jpg
jpeg(filename = "figures/Taylor_Valley_response_ratio_interaction_plot.jpeg", 
     width = 1400, 
     height = 600, 
     units = "px", 
     pointsize = 14, 
     quality = 100, 
     bg = "white", 
     res = NA)
#
fg3.all
#
# Close device
dev.off()
```
