---
title: "Nutrient Limitation in the MDV of Antarctica"
subtitle: "  \n Logistic Growth Model - Inter-basin"
author: "Adam J. Solon"
date: "`r Sys.Date()`"
#output: html_document
output: 
  pdf_document:
    #includes:  
     # in_header: my_header.txt
    toc: TRUE
    fig_width: 8
    fig_height: 6
    fig_caption: true
fontsize: 12pt
#editor_options: 
#  chunk_output_type: console
---

# Script Summary  
This report quantifies the growth rate and carrying capacity of primary producers from MDV soils using a logistic growth model.  The experiment involved soils from Taylor Valley in the McMurdo Dry Valleys (MDV) of Antarctica and were collected from mountain slopes in one of three lake basins: Bonney(interior), Hoare(intermediate), and Fryxell(coastal).  There are four treatments: CTRL (controls with H2O only), N (nitrogen + H2O only), P (phosphorus + H2O only), and NP (N and P combined + H2O). 

### Steps of this pipeline:  
1.  Create and organize directories
2.  Load R packages
3.  Subset according to basin
4.  Input files
5.  Format Files
6.  Run analyses
7.  Plot curves  
8.  Statistics  

```{r, echo = FALSE, include = FALSE}
# Change identifiers to your system and file naming. 
user <- "C:/Users/adams" # computer user name
folder <- "/Desktop/Projects"
project <- "/MDV_Soils_nutrient_limitation/manuscript1"
analyses <- "/analyses"
data  <- "/point_intercept"
analysis4 <- "/GR_growth_rate"
ecosystem <- "lab microcosms - Taylor Valley soils" # Define the environment for use in report language.
project.fp <- paste0(user, folder, project)
```

# Pipeline

### Set paths and create directories 

```{r set paths for project, echo = FALSE, include = FALSE}
# First define the project and project folders. 

# Create project directory
###################################################
project.fp <- paste0(user, folder, project)
if (!dir.exists(project.fp)) dir.create(project.fp)

# Create sub-directory for all analyses for project 
###################################################
analyses.fp <- paste0(project.fp, analyses)
if (!dir.exists(analyses.fp)) dir.create(analyses.fp)

# Create sub-directory for data type 
###################################################
data.fp <- paste0(analyses.fp, data)
if (!dir.exists(data.fp)) dir.create(data.fp)

# Create sub-directory for specific analysis 
###################################################
analysis4.fp <- paste0(data.fp, analysis4)
if (!dir.exists(analysis4.fp)) dir.create(analysis4.fp)

# Create sub-directories for analysis 
###################################################

# Create directory for 'before' pipeline inputs
input.fp <- paste0(analysis4.fp, "/input")
if (!dir.exists(input.fp)) dir.create(input.fp)

# Create directory for 'within' pipeline R objects 
objects.fp <- paste0(analysis4.fp, "/objects")
if (!dir.exists(objects.fp)) dir.create(objects.fp)

# Create directory for 'after' pipeline outputs 
output.fp <- paste0(analysis4.fp, "/output")
if (!dir.exists(output.fp)) dir.create(output.fp)

# Check for the folders here as a sanity check. Should see "Input" and "Objects" if starting from scratch.
list.files(analysis4.fp) 

```

### Create additional sub-directories by analysis and basin  
  
```{r echo = FALSE, include = FALSE}
# Create sub-directory for Growth Rates (GR) analysis
###################################################
analysis4.fp <- paste0(data.fp, analysis4)
if (!dir.exists(analysis4.fp)) dir.create(analysis4.fp)

# Create sub-directories for GR 
###################################################

# Create directory for 'before' pipeline inputs
input4.fp <- paste0(analysis4.fp, "/input")
if (!dir.exists(input4.fp)) dir.create(input4.fp)

# Create directory for 'within' pipeline R objects 
objects4.fp <- paste0(analysis4.fp, "/objects")
if (!dir.exists(objects4.fp)) dir.create(objects4.fp)

# Create directory for 'after' pipeline outputs 
output4.fp <- paste0(analysis4.fp, "/output")
if (!dir.exists(output4.fp)) dir.create(output4.fp)

# Check for the folders here as a sanity check. Should see "Input" and "Objects" if starting from scratch.
list.files(analysis4.fp) 

# Create sub-directories for Bonney basin 
###################################################

# Create directory for 'within' pipeline R objects 
objects4.b.fp <- paste0(objects4.fp, "/bonney")
if (!dir.exists(objects4.b.fp)) dir.create(objects4.b.fp)

# Create sub-directories for Hoare basin 
###################################################

# Create directory for 'within' pipeline R objects 
objects4.h.fp <- paste0(objects4.fp, "/hoare")
if (!dir.exists(objects4.h.fp)) dir.create(objects4.h.fp)

# Create sub-directories for Fryxell basin 
###################################################

# Create directory for 'within' pipeline R objects 
objects4.f.fp <- paste0(objects4.fp, "/fryxell")
if (!dir.exists(objects4.f.fp)) dir.create(objects4.f.fp)

```

### Session Information  

```{r Install and load packages, echo = FALSE, include = FALSE}

# install.packages("tidyverse")
# install.packages("knitr")
# install.packages("kableExtra")
# install.packages("lmtest")
# install.packages("AICcmodavg")
# install.packages("emmeans")

library(tidyverse); packageVersion("tidyverse")
library(knitr); packageVersion("knitr")
library(kableExtra); packageVersion("kableExtra")
library(lmtest); packageVersion("lmtest")
library(AICcmodavg); packageVersion("AICcmodavg")
library(emmeans); packageVersion("emmeans")

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

```  
  
* r version: `r getRversion()`
* RStudio version: `r rstudioapi::versionInfo()$version`
* r packages:  
  tidyverse, `r packageVersion("tidyverse")` 
  knitr, `r packageVersion("knitr")`  
  kableExtra, `r packageVersion("KableExtra")`  
  lmtest, `r packageVersion("lmtest")`   
  AICcmodavg, `r packageVersion("AICcmodavg")`  
  emmeans, `r packageVersion("emmeans")`  

### Input files 
1. Growth models from each basin, filenames 'FXM_grow_mod_params.rds', 'HRM_grow_mod_params.rds', 'BNM_grow_mod_params.rds' from scripts MCosms_II_growth_model_fryxell.Rmd, MCosms_II_growth_model_hoare.Rmd, and MCosms_II_growth_model_bonney.Rmd

```{r input data files, echo = FALSE, include = FALSE}
# input data files
# Count data from point-intercept (microscopy)
gc.b.fp <- paste0(objects4.b.fp, "/BNM_grow_mod_params.rds") 
gc.h.fp <- paste0(objects4.h.fp, "/HRM_grow_mod_params.rds") 
gc.f.fp <- paste0(objects4.f.fp, "/FXM_grow_mod_params.rds") 

#input file
gc.b <- readRDS(gc.b.fp)
gc.h <- readRDS(gc.h.fp)
gc.f <- readRDS(gc.f.fp)

```
  
### Format files  
1. Add column with basin for each data frame
2. Bind data frames by rows
3. Subset by treatment- P or NP

```{r format all data, echo = FALSE, include = FALSE}
#add basin column to data frames
gc.b$basin <- "bonney"
gc.h$basin <- "hoare"
gc.f$basin <- "fryxell"

#merge data frames
gc.all <- rbind(gc.b, gc.h, gc.f)

#treatment as factor
gc.all$treatment <- as.factor(gc.all$treatment) 

#subset by P treatment
gc.all.p <- gc.all %>% filter(treatment %in% "P")

#subset by NP treatment
gc.all.np <- gc.all %>% filter(treatment %in% "NP")


```
  
## Run analyses
  
### Linear Model - Maximum growth rate (mumax) - NP treatments  
  
```{r test for differences in NP treatment growth rates, echo=FALSE}
#lm - growth rates (mumax) for NP treatments across basins
lm.np.mumax <- lm(mumax ~ basin,  gc.all.np)

#output from model
summary(lm.np.mumax)

#diagnostics
plot(lm.np.mumax)

```
  
```{r NP null model, echo=FALSE}
#null model
null.np.mumax <- lm(mumax ~ 1,  gc.all.np)

#output from null model
summary(null.np.mumax)

```

```{r test with NP null model, echo=FALSE}
#test significance relative to null model that assumes no treatment effect
lrtest(lm.np.mumax, null.np.mumax) ##likelihood-ratio test significant difference betwen null model and treatment model. 
```
  
```{r NP model comparison - mumax, echo=FALSE, include = FALSE}
#lm
#second order AICc (small sample size)
aicC.np.mumax <- AICc(lm.np.mumax, return.K = FALSE, second.ord = TRUE) 

#display
print(aicC.np.mumax)

#parameters in model
aicC.np.mumax.p <- AICc(lm.np.mumax, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.np.mumax.p)

#null
#second order AICc (small sample size)
aicC.null.mumax <- AICc(null.np.mumax, return.K = FALSE, second.ord = TRUE) 

#display
print(aicC.null.mumax)

#parameters in model
aicC.null.mumax.p <- AICc(null.np.mumax, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.null.mumax.p)
```
  
```{r echo = FALSE, include = FALSE}
#create data frame w/ AICc results
aicC.m <- data.frame (Model  = c("lm", "lm.null"),
                  AICc = c(aicC.np.mumax, aicC.null.mumax),
                  parameters = c(aicC.np.mumax.p, aicC.null.mumax.p)
                  )
```
  
```{r kable 1}
#create table of AIC values
knitr::kable(aicC.m, booktabs = T, digits = 2, align = "cc", caption = 'Model Comparison with Akaike Information Criterion (AIC)') %>%
  kable_styling(latex_options = c("striped", "hold_position"), font_size = 14) %>%
  row_spec(0, bold = T)

```
  
```{r NP fitted model pairwise comparison, echo=FALSE}
#save as .txt file
capture.output(summary(lm.np.mumax), file = paste0(output4.fp, '/Inter_basin_NP_mumax_lm_summary.txt'))

#post-hoc test of fitted models
p.hoc.mumax.np <- test(pairs(emmeans(lm.np.mumax, ~ basin, mode = "link")))

#output
print(p.hoc.mumax.np)

#save as .csv file
write.csv(p.hoc.mumax.np, paste0(output4.fp, file = '/Inter_basin_NP_mumax_pairwise.csv'))

```
  
### Linear Model - Carrying Capacity (K) - NP treatments  
  
```{r test for differences in NP treatment K, echo=FALSE}
#lm - growth rates (mumax) for NP treatments across basins
lm.np.K <- lm(K ~ basin,  gc.all.np)

#output from model
summary(lm.np.K)

#diagnostics
plot(lm.np.K)

```
  
```{r NP null model K, echo=FALSE}
#null model
null.np.K <- lm(K ~ 1,  gc.all.np)

#output from null model
summary(null.np.K)

```

```{r test with NP null model K, echo=FALSE}
#test significance relative to null model that assumes no treatment effect
lrtest(lm.np.K, null.np.K) ##likelihood-ratio test significant difference betwen null model and treatment model. 
```
  
```{r NP model comparison K, echo = FALSE, include = FALSE}
#lm
#second order AICc (small sample size)
aicC.np.K <- AICc(lm.np.K, return.K = FALSE, second.ord = TRUE) 

#display
print(aicC.np.K)

#parameters in model
aicC.np.K.p <- AICc(lm.np.K, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.np.K.p)

#null
#second order AICc (small sample size)
aicC.null.K <- AICc(null.np.K, return.K = FALSE, second.ord = TRUE) 

#display
print(aicC.null.K)

#parameters in model
aicC.null.K.p <- AICc(null.np.K, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.null.K.p)
```
  
```{r echo = FALSE, include = FALSE}
#create data frame w/ AICc results
aicC.K <- data.frame (Model  = c("lm", "lm.null"),
                  AICc = c(aicC.np.K, aicC.null.K),
                  parameters = c(aicC.np.K.p, aicC.null.K.p)
                  )
```
  
```{r kable 2}
#create table of AIC values
knitr::kable(aicC.K, booktabs = T, digits = 2, align = "cc", caption = 'Model Comparison with Akaike Information Criterion (AIC)') %>%
  kable_styling(latex_options = c("striped", "hold_position"), font_size = 14) %>%
  row_spec(0, bold = T)

```
  
```{r fitted model pairwise, echo=FALSE}
#save as .txt file
capture.output(summary(lm.np.K), file =  paste0(output4.fp, '/Inter_basin_NP_K_lm_summary.txt'))

#post-hoc test of fitted models
p.hoc.K.np <- test(pairs(emmeans(lm.np.K, ~ basin, mode = "link")))

#output
print(p.hoc.K.np)

#save as .csv file
write.csv(p.hoc.K.np, file = paste0(output4.fp, '/Inter_basin_NP_K_pairwise.csv'))

```
  
### Linear Model - Maximum growth rate (mumax) - P treatments  
  
```{r test for differences in P treatment growth rates, echo=FALSE}
#lm - growth rates (mumax) for P treatments across basins
lm.p.mumax <- lm(mumax ~ basin,  gc.all.p)

#output from model
summary(lm.p.mumax)

#diagnostics
plot(lm.p.mumax)

```
  
```{r null model, echo=FALSE}
#null model
null.p.mumax <- lm(mumax ~ 1,  gc.all.p)

#output from null model
summary(null.p.mumax)

```

```{r test with null model, echo=FALSE}
#test significance relative to null model that assumes no treatment effect
lrtest(lm.p.mumax, null.p.mumax) ##likelihood-ratio test significant difference betwen null model and treatment model. 
```
  
```{r model comparison, echo=FALSE}
#lm
#second order AICc (small sample size)
aicC.p.mumax <- AICc(lm.p.mumax, return.K = FALSE, second.ord = TRUE) 

#display
print(aicC.p.mumax)

#parameters in model
aicC.p.mumax.p <- AICc(lm.p.mumax, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.p.mumax.p)

#null
#second order AICc (small sample size)
aicC.null.mumax <- AICc(null.p.mumax, return.K = FALSE, second.ord = TRUE) 

#display
print(aicC.null.mumax)

#parameters in model
aicC.null.mumax.p <- AICc(null.p.mumax, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.null.mumax.p)
```
  
```{r echo = FALSE, include = FALSE}
#create data frame w/ AICc results
aicC.m <- data.frame (Model  = c("lm", "lm.null"),
                  AICc = c(aicC.p.mumax, aicC.null.mumax),
                  parameters = c(aicC.p.mumax.p, aicC.null.mumax.p)
                  )
```
  
```{r kable 3}
#create table of AIC values
knitr::kable(aicC.m, booktabs = T, digits = 2, align = "cc", caption = 'Model Comparison with Akaike Information Criterion (AIC)') %>%
  kable_styling(latex_options = c("striped", "hold_position"), font_size = 14) %>%
  row_spec(0, bold = T)

```
  
```{r P fitted model pairwise, echo=FALSE, include = FALSE}
#save as .txt file
capture.output(summary(lm.p.mumax), file = paste0(output4.fp, '/Inter_basin_P_mumax_lm_summary.txt'))

```
  
### Linear Model - Carrying Capacity (K) - P treatments  
  
```{r test for differences in P treatment K, echo=FALSE}
#lm - growth rates (mumax) for P treatments across basins
lm.p.K <- lm(K ~ basin,  gc.all.p)

#output from model
summary(lm.p.K)

#diagnostics
plot(lm.p.K)

```
  
```{r P null model K, echo=FALSE}
#null model
null.p.K <- lm(K ~ 1,  gc.all.p)

#output from null model
summary(null.p.K)

```

```{r P test with null model K, echo=FALSE}
#test significance relative to null model that assumes no treatment effect
lrtest(lm.p.K, null.p.K) ##likelihood-ratio test significant difference betwen null model and treatment model. 
```
  
```{r P model comparison K, echo=FALSE}
#lm
#second order AICc (small sample size)
aicC.p.K <- AICc(lm.p.K, return.K = FALSE, second.ord = TRUE) 

#display
print(aicC.p.K)

#parameters in model
aicC.p.K.p <- AICc(lm.p.K, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.p.K.p)

#null
#second order AICc (small sample size)
aicC.null.K <- AICc(null.p.K, return.K = FALSE, second.ord = TRUE) 

#display
print(aicC.null.K)

#parameters in model
aicC.null.K.p <- AICc(null.p.K, return.K = TRUE, second.ord = TRUE)

#display
print(aicC.null.K.p)
```
  
```{r echo = FALSE, include = FALSE}
#create data frame w/ AICc results
aicC.K <- data.frame (Model  = c("lm", "lm.null"),
                  AICc = c(aicC.p.K, aicC.null.K),
                  parameters = c(aicC.p.K.p, aicC.null.K.p)
                  )
```
  
```{r kable 4}
#create table of AIC values
knitr::kable(aicC.K, booktabs = T, digits = 2, align = "cc", caption = 'Model Comparison with Akaike Information Criterion (AIC)') %>%
  kable_styling(latex_options = c("striped", "hold_position"), font_size = 14) %>%
  row_spec(0, bold = T)

```
  
```{r P fitted model pairwise K, echo=FALSE, include = FALSE}
#save as .txt file
capture.output(summary(lm.p.K), file =  paste0(output4.fp, '/Inter_basin_P_K_lm_summary.txt'))

```